### 2020-02-17 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-Title-Page-v1.Rmd

---
title: An empirical assessment of strategies to model opponent effects in crash severity analysis
author:
  - name: Antonio Paez
    email: paezha@mcmaster.ca
    affiliation: McMaster University
    footnote: Corresponding Author
  - name: Hany Hassan
    email: hassan1@lsu.edu
    affiliation: Louisiana State University
  - name: Mark Ferguson
    email: fergumr@mcmaster.ca
    affiliation: McMaster University
  - name: Saiedeh Razavi
    email: razavi@mcmaster.ca
    affiliation: McMaster University    
address:
  - code: McMaster University
    address: McMaster Institute for Transportation and Logistics, McMaster University, 1280 Main Street West, Hamilton, Ontario, Canada L8S 4K1
  - code: Louisiana State University
    address: Department of Civil and Environmental Engineering, Louisiana State University, Baton Rouge, Louisiana, USA 70803
abstract: |
  Road accidents impose an important burden on health and the economy. Numerous efforts have been undertaken to understand the factors that affect road collisions. One stream of research is on modelling the severity of crashes. In this literature several strategies have been proposed to deal with the way the different parties in a crash interact to influence the severity of the individual outcomes, including the use of opponent variables, data subsetting, and estimating models with hierarchical components. However, no systematic assessment has been conducted of the performance of these strategies. The objective of this paper is to empirically evaluate these modelling strategies in the context of crashes involving two parties. To this end, a series of models are estimated using data from Canada's National Collision Database. Three levels of crash severity (no injury/injury/fatality) are analyzed using ordered logit models and covariates for the participants in the crash and the conditions of the crash. The models are assessed using predicted shares and classes of outcomes, and the results highlight the importance of considering opponent effects in crash severity analysis. The study also suggests that hierarchical (i.e., multi-level) specifications and subsetting do not necessarily perform better than a relatively simple single-level model with opponent effects. The results of this study provide insights regarding the performance of different modelling strategies, and should be informative to researchers working with crash severity models.

journal: "Some Journal"
date: "`r Sys.Date()`"
bibliography: mybibfile.bib
#linenumbers: true
numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article

header-includes:
   - \usepackage[margin=1in]{geometry}
   - \usepackage{lineno}
   - \linenumbers
---