### 2020-05-10 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-Title-Page-v2.tex

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Some Journal} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={An empirical assessment of strategies to model party interactions in crash severity analysis},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc header
\usepackage[margin=1in]{geometry}
\usepackage{lineno}
\linenumbers

\begin{document}
\begin{frontmatter}

  \title{An empirical assessment of strategies to model party interactions in
crash severity analysis}
    \author[McMaster University]{Antonio Paez\corref{Corresponding Author}}
   \ead{paezha@mcmaster.ca} 
    \author[Louisiana State University]{Hany Hassan}
   \ead{hassan1@lsu.edu} 
    \author[McMaster University]{Mark Ferguson}
   \ead{fergumr@mcmaster.ca} 
    \author[McMaster University]{Saiedeh Razavi}
   \ead{razavi@mcmaster.ca} 
      \address[McMaster University]{McMaster Institute for Transportation and Logistics, McMaster
University, 1280 Main Street West, Hamilton, Ontario, Canada L8S 4K1}
    \address[Louisiana State University]{Department of Civil and Environmental Engineering, Louisiana State
University, Baton Rouge, Louisiana, USA 70803}

  \begin{abstract}
  Road crashes impose an important burden on health and the economy.
  Numerous efforts have been undertaken to understand the factors that
  affect road collisions in general, and the severity of crashes in
  particular. In this literature several strategies have been proposed to
  model interactions between parties in a crash, including the use of
  variables regarding the other party (or parties) in the collision, data
  subsetting, and estimating models with hierarchical components. Since no
  systematic assessment has been conducted of the performance of these
  strategies, they appear to be used in an ad-hoc fashion in the
  literature. The objective of this paper is to empirically evaluate ways
  to model party interactions in the context of crashes involving two
  parties. To this end, a series of models are estimated using data from
  Canada's National Collision Database. Three levels of crash severity (no
  injury/injury/fatality) are analyzed using ordered probit models and
  covariates for the parties in the crash and the conditions of the crash.
  The models are assessed using predicted shares and classes of outcomes,
  and the results highlight the importance of considering opponent effects
  in crash severity analysis. The study also suggests that hierarchical
  (i.e., multi-level) specifications and subsetting do not necessarily
  perform better than a relatively simple single-level model with
  opponent-related factors. The results of this study provide insights
  regarding the performance of different modelling strategies, and should
  be informative to researchers in the field of crash severity.
  \end{abstract}

 \end{frontmatter}

\end{document}