### 2020-05-10 Interactions-Crash-Severity-v1/Interactions-Crash-Severity-v2.tex

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Accident Analysis and Prevention} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={An empirical assessment of strategies to model party interactions in crash severity analysis},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc header
\usepackage[margin=1in]{geometry}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\doublespacing
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\begin{frontmatter}

  \title{An empirical assessment of strategies to model party interactions in
crash severity analysis}
    \author[McMaster University]{Antonio Paez\corref{Corresponding Author}}
   \ead{paezha@mcmaster.ca} 
    \author[Louisiana State University]{Hany Hassan}
   \ead{hassan1@lsu.edu} 
    \author[McMaster University]{Mark Ferguson}
   \ead{fergumr@mcmaster.ca} 
    \author[McMaster University]{Saiedeh Razavi}
   \ead{razavi@mcmaster.ca} 
      \address[McMaster University]{McMaster Institute for Transportation and Logistics, McMaster
University, 1280 Main Street West, Hamilton, Ontario, Canada L8S 4K1}
    \address[Louisiana State University]{Department of Civil and Environmental Engineering, Louisiana State
University, Baton Rouge, Louisiana, USA 70803}

  \begin{abstract}
  Road crashes impose an important burden on health and the economy.
  Numerous efforts have been undertaken to understand the factors that
  affect road collisions in general, and the severity of crashes in
  particular. In this literature several strategies have been proposed to
  model interactions between parties in a crash, including the use of
  variables regarding the other party (or parties) in the collision, data
  subsetting, and estimating models with hierarchical components. Since no
  systematic assessment has been conducted of the performance of these
  strategies, they appear to be used in an ad-hoc fashion in the
  literature. The objective of this paper is to empirically evaluate ways
  to model party interactions in the context of crashes involving two
  parties. To this end, a series of models are estimated using data from
  Canada's National Collision Database. Three levels of crash severity (no
  injury/injury/fatality) are analyzed using ordered probit models and
  covariates for the parties in the crash and the conditions of the crash.
  The models are assessed using predicted shares and classes of outcomes,
  and the results highlight the importance of considering opponent effects
  in crash severity analysis. The study also suggests that hierarchical
  (i.e., multi-level) specifications and subsetting do not necessarily
  perform better than a relatively simple single-level model with
  opponent-related factors. The results of this study provide insights
  regarding the performance of different modelling strategies, and should
  be informative to researchers in the field of crash severity.
  \end{abstract}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Modelling the severity of injuries to victims of road crashes has been a
preoccupation of transportation researchers, planners, auto insurance
companies, governments, and the general public for decades. One of the
earliest studies to investigate the severity of injuries conditional on
a collision having occurred was by White and Clayton (1972). Kim et
al.~(1995) later stated that the ``linkages between severity of injury
and driver characteristics and behaviors have not been thoroughly
investigated'' (p.~470). Nowadays, there is a burgeoning literature on
this subject, which often relies on multivariate analysis of crash
severity to clarify the way various factors can affect the outcome of an
incident, and to discriminate between various levels of injury.

Crash severity is an active area of research, and one where
methodological developments have aimed at improving the reliability,
accuracy, and precision of models (e.g., Savolainen et al., 2011; Bogue
et al., 2017; Yasmin and Eluru, 2013). Of interest in this literature is
how different parties in a crash interact to influence the severity of
individual outcomes. The importance of these interactions has been
recognized by numerous authors (e.g., Chiou et al., 2013; Lee and Li,
2014; Li et al., 2017; Torrao et al., 2014). Lee and Li (2014) for
instance, note that ``{[}for{]} two-vehicle crashes, most studies only
considered the effects of one vehicle on driver's injury severity or the
highest injury severity of two drivers. However, it is expected that
driver's injury severity is not only affected by characteristics of
his/her own vehicle, but also characteristics of a partner vehicle.''
More generally, the severity of the outcome depends, at least in part,
on the characteristics of the parties, and a crash between two heavy
vehicles is likely to have very different outcomes compared to crash
where a heavy vehicle hits a pedestrian.

For the purpose of this paper, we define a party as one or more
individuals travelling in a traffic unit that becomes involved in a
crash. Sometimes the traffic unit is a vehicle, and the party is a
single individual (i.e.~a single occupant vehicle); but in other cases,
a party could consist of several individuals (i.e., a driver and one or
more passengers). Other times, the individual \emph{is} the traffic
unit, for instance a pedestrian or a bicyclist. An opponent is a
different party that is involved in the same collision. In the
literature, interactions between parties in a collision are treated by
means of different strategies, including the use of \emph{data
subsetting}, \emph{opponent variables}, and estimating models with
\emph{hierarchical components}. These strategies are not new, however, a
systematic comparison between them is missing from the literature. For
this reason, the objective of this paper is to empirically evaluate
different strategies to model party interactions in crash severity in
the context of incidents involving two parties. More concretely, this
research aims to:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Systematize the analysis of party interactions in crash severity
  models. Although every strategy considered here has been used in past
  research, in this paper they are organized in a way that clarifies
  their operation.
\item
  Present a data management workflow to prepare a dataset to implement
  analysis of opponent effects.
\item
  Provide evidence of the performance of different modelling strategies.
  In particular, the importance of considering opponent-level effects
  and the suitability of single-level models.
\item
  Present an example of reproducible research in crash severity
  analysis: all data and code are publicly available from the beginning
  of the peer-review
  process\footnote{The source file for this paper is an `R` Markdown document; all code and data necessary to reproduce the analysis are available from the following GitHub repository: \url{https://github.com/paezha/Modelling-Participant-Interactions-in-Crash-Severity}}.
\end{enumerate}

For the assessment we use data from Canada's National Collision
Database, a database that collects all police-reported collisions in the
country. Using the most recent version of the data set (2017). Three
levels of crash severity (no injury/injury/fatality) are analyzed using
ordered logit models, and covariates for the parties in the crash and
the conditions of the crash. For model assessment, we conduct an
in-sample prediction exercise using the estimation sample (i.e.,
\emph{nowcasting}), and also an out-of-sample prediction exercise using
the data set corresponding to 2016 (i.e., \emph{backcasting}). The
models are assessed using predicted shares and predicted classes of
outcomes at the individual level, using an extensive array of
verification statistics.

The rest of this paper is structured as follows. In Section
\ref{sec:background} we discuss some background matters, and follow this
with a concise review of the modelling strategies used to analyze crash
severity \ref{sec:methods}. Section \ref{sec:application} describes the
data requirements, data preprocessing, and the modelling strategies,
along with the results of model estimation. The results of assessing the
models and the discussion of these results is found in Section
\ref{sec:assessment}. We then present some additional thoughts about the
applicability of this approach in Section
\ref{sec:further-considerations} before offering some concluding remarks
in Section \ref{sec:concluding-remarks}.

\hypertarget{sec:background}{%
\section{Background}\label{sec:background}}

Crash severity is often modeled using models for discrete outcomes.
Analysts interested in crash severity have at their disposal an ample
repertoire of models to choose from, including classification techniques
from machine learning (e.g., Iranitalab and Khattak, 2017; Chang and
Wang, 2006; Effati et al., 2015; Khan et al., 2015), Poisson models for
counts (e.g., Ma et al., 2008), unordered logit/probit models (e.g., Tay
et al., 2011), as well as ordered logit/probit models (e.g., Rifaat and
Chin, 2007), with numerous variants, such as random parameters/mixed
logit (e.g., Aziz et al., 2013; Haleem and Gan, 2013), partial
proportional odds models (e.g., Mooradian et al., 2013; Sasidharan and
Menendez, 2014), and the use of copulas (e.g., Wang et al., 2015).
Recent reviews of methods include Savolainen et al.~(2011), Yasmin and
Eluru (2013), and Mannering et al.~(2016).

Irrespective of the modelling framework employed, models of crash
severity often include variables in several categories, as shown with
examples in Table \ref{tab:variable-categories} (also see Montella et
al., 2013). Many crash databases and analyses also account for the
multi-event nature of many crashes. Individuals in the crash may have
had different roles depending on their situation, with some acting as
operators of a vehicle (i.e., drivers, bicyclists), while others were
passengers. They also may differ depending on what type of traffic unit
they were, for example pedestrians, or operators/passengers of a
vehicle. The multiplicity of roles makes for complicated modelling
decisions when trying to understand the severity of injuries; for
example, what is the unit of analysis, the person, the traffic unit, or
the collision? Not surprisingly, it is possible to find examples of
studies that adopt different perspectives. A common simplifying strategy
in model specification is to consider only \emph{drivers} and/or only
\emph{single-vehicle} crashes (e.g., Kim et al., 2013; Gong and Fan,
2017; Lee and Li, 2014; Osman et al., 2018). This strategy reduces the
dimensions of the event, and it becomes possible, for example, to equate
the traffic unit to the person for modelling purposes.

The situation becomes more complex when dealing with events that involve
two traffic units (e.g., Torrao et al., 2014; Wang et al., 2015) and
multi-traffic unit crashes (e.g., Wu et al., 2014; Bogue et al., 2017).
Different strategies have been developed to study these, more complex
events. A number of studies advocate the estimation of separate models
for different parties, individuals, and/or situations. In this way, Wang
and Kockelman (2005) estimated models for single-vehicle and two-vehicle
crashes, while Savolainen and Mannering (2007) estimated models for
single-vehicle and multi-vehicle crashes. More recently, Duddu et
al.~(2018) and Penmetsa et al.~(2017) presented research that estimated
separate models for at-fault and not-at-fault drivers. The strategy of
estimating separate models relies on \emph{subsetting} the data set,
although it is possible to link the relevant models more tightly by
means of a common covariance structure, as is the case of bivariate
models (e.g., Chiou et al., 2013; Chen et al., 2019) or models with
copulas (e.g., Rana et al., 2010; Shamsunnahar et al., 2014; Wang et
al., 2015).

A related strategy to specify crash severity models is to organize the
data in such a way that it is possible to introduce \emph{opponent
effects}. There are numerous examples of studies that consider at least
some characteristics of the opposite party (or parties) in two- or
multi-vehicle crashes. For example, Wang and Kockelman (2005) considered
the type of the opposite vehicle in their model for two-vehicle
collisions. Similarly, Torrao et al.~(2014) included in their analysis
the age, wheelbase, weight, and engine size of the opposite vehicle,
while Bogue et al.~(2017) used the body type of the opposite vehicle.
Penmetsa et al.~(2017) and Duddu et al.~(2018) are two of the most
comprehensive examples of using opponent's information, as they included
attributes of individuals in the opposite party (their physical
condition, sex, and age), as well as characteristics of the other
party's traffic unit (the vehicle type of the opponent). The twin
strategies of subsetting the sample and using the attributes of the
opponent are not mutually exclusive, but neither are they used
consistently together, as a scan of the literature reveals.

Another strategy is to introduce \emph{hierarchical components} in the
model, a technique widely used in the hierarchical or multi-level
modelling literature. This involves considering observations as being
nested at different levels: individuals nested in traffic units, which
in turn are nested in accidents, as an example.

In this paper we consider three general modelling strategies, as
follows:

\begin{itemize}
\item
  Strategy 1. Introducing opponent-related factors
\item
  Strategy 2. Single-level model and multi-level (hierarchical) model
  specifications
\item
  Strategy 3. Full sample and sample subsetting
\end{itemize}

These are discussed in more detail in the following section.

\begin{table}

\caption{\label{tab:table-variable-categories}\label{tab:variable-categories}Categories of variables used in the analysis of crash severity with examples}
\centering
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{22em}}
\toprule
Category & Examples\\
\midrule
Human-factors & Attributes of individuals in the crash, e.g., injury status, age, gender, licensing status, professional driver status\\
Traffic unit-factors & Attributes of the traffic unit, e.g., type of traffic unit (pedestrian, car, motorcycle, etc.), maneuver, etc.\\
Environmental-factors & Attributes of the crash, e.g., location, weather conditions, light conditions, number of parties, etc.\\
Road-factors & Attributes of the road, e.g., surface condition, grade, geometry, etc.\\
Opponent-related factors & Attributes of the opponent, e.g., age of opponent, gender of opponent, opponent vehicle type, etc.\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{sec:methods}{%
\section{Methods}\label{sec:methods}}

\hypertarget{choice-of-model}{%
\subsection{Choice of model}\label{choice-of-model}}

Before describing the modelling strategies, it is important to explain
our choice of model. There have already been comparisons between
different models. Yasmin and Eluru (2013), for instance, conducted an
extensive comparison of models for discrete outcomes in crash severity
analysis, and found only small differences in the performance of
unordered models and ordered models; however, ordered models are usually
more parsimonious since only one latent functions needs to be estimated
for all outcomes, as opposed to one for each outcome in unordered
models. Bogue et al.~(2017) also compared unordered and ordered models
in the form of the mixed multinomial logit and a modified rank ordered
logit, respectively, and found that the ordered model performed best. To
keep the empirical assessment manageable, in this paper we will consider
only the ordinal logit model, and will comment on potential extensions
in Section \ref{sec:further-considerations}.

The ordinal model is a latent-variable approach, whereby the severity of
the crash (observed) is linked to an underlying latent variable that is
a function of the variables of interest, as follows:

\begin{equation}
\label{eq:latent-function}
y_{itk}^*=\sum_{l=1}^L\alpha_lp_{itkl} + \sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \epsilon_{itk}
\end{equation}

The left-hand side of the expression above (\(y_{itk}^*\)) is a latent
(unobservable) variable that is associated with the severity of crash
\(k\) (\(k=1,\cdots,K\)) for individual \(i\) in traffic unit \(t\). The
right-hand side of the expression is split in four parts. The first part
gathers \(l=1,\cdots,L\) human-factors \(p\) for individual \(i\) in
traffic unit \(t\) and crash \(k\); these could relate to the person
(e.g., age, gender, and road user class). The second part gathers
\(m=1,\cdots,M\) attributes \(u\) related to traffic unit \(t\) in crash
\(k\); these could be items such as maneuver or vehicle type. The third
part gathers \(q=1,\cdots,Q\) attributes \(c\) related to the crash
\(k\), including environmental-factors and road-factors, such as weather
conditions, road alignment, and type of surface. Lastly, the fourth
element is a random term specific to individual \(i\) in traffic unit
\(t\) and crash \(k\). The function consists of a total of \(Z=L+M+Q\)
covariates and associated parameters.

For conciseness, in what follows we will abbreviate the function as
follows:

\begin{equation}
\label{eq:latent-function-compact}
y_{itk}^*=\sum_{z=1}^Z\theta_zx_{itkz} + \epsilon_{itk}
\end{equation}

The latent variable is not observed directly, but it is possible to
posit a probabilistic relationship with the outcome \(y_{itk}\) (the
severity of crash \(k\) for individual \(i\) in traffic unit \(t\)).
Depending on the characteristics of the data and the assumptions made
about the random component of the latent function different models can
be obtained. For example, if crash severity is coded as a variable with
three outcomes (e.g., property damage only/injury/fatal), we can relate
the latent variable to the outcome as follows:

\begin{equation}
\label{eq:latent-function-ordered-outcomes}
y_{itk} = 
\begin{cases}
\text{fatality} & \text{if } y_{itk}^*> \mu_2\\
\text{injury} & \text{if } \mu_1< y_{itk}^*< \mu_2\\
\text{PDO} & \text{if } y_{itk}^*< \mu_1
\end{cases}
\end{equation}

\noindent where \(\mu_1\) and \(\mu_2\) are estimable thresholds. Due to
the stochastic nature of the latent function, the outcome of the crash
is not fully determined. However, we can make the following probability
statements:

\begin{equation}
\label{eq:probability-ordered-outcomes}
\begin{array}{rcl}\
P(y_{itk} = \text{PDO}) &=& 1 - P(y_{itk} = \text{injury}) - P(y_{itk} = \text{fatality})\\ 
P(y_{itk} = \text{injury}) &=& P(\mu_1 - \sum_{z=1}\theta_zp_{itkz} < \epsilon_{itk} < \mu_2 - \sum_{z=1}\theta_zp_{itkz})\\
P(y_{itk} = \text{fatality}) &=& P(\epsilon_{itk} < \mu_1 - \sum_{z=1}\theta_zp_{itkz})
\end{array}
\end{equation}

If the random terms are assumed to follow the logistic distribution, the
ordered logit model is obtained; if the normal distribution, then the
ordered probit model. Estimation methods for these models are very
well-established (e.g., Maddala, 1986; Train, 2009). There are numerous
variations of the basic modelling framework above, including
hierarchical models, bivariate models, multinomial models, and Bayesian
models, among others (see Savolainen et al., 2011 for a review of
methods).

\hypertarget{strategy-1-opponent-related-factors}{%
\subsection{Strategy 1: opponent-related
factors}\label{strategy-1-opponent-related-factors}}

When opponent-related variables are included, the function is augmented
as follows:

\begin{equation}
\label{eq:latent-function-with-opponent-variables}
y_{itk}^*=\sum_{l=1}^L\alpha_lp_{itkl} + \sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \sum_{r=1}^R\delta_ro_{jvkr} + \epsilon_{itk}
\end{equation}

This function includes one additional summation compared to Equation
\ref{eq:latent-function-ordered-outcomes}. This summation gathers
\(r=1,\cdots,R\) attributes \(o\) related to individual \(j\) in traffic
unit \(v\) that was an opposite party to individual \(i\) in traffic
unit \(t\) in crash \(k\). These attributes could be individual
characteristics of the operator of the opposite traffic unit (such as
age and gender) and/or characteristics of the opposite traffic unit
(such as vehicle type or weight). To qualify as an opposite party,
individual \(j\) must have been an individual in crash \(k\) but
operating traffic unit \(v\ne t\). Sometimes the person \emph{is} the
traffic unit, as is the case of a pedestrian. And we exclude passengers
of vehicles as opponents, since they do not operate the traffic unit. In
case the opponent attributes include only characteristics of the traffic
unit, the condition for the traffic unit to be an opponent is that it
was part of crash \(k\) and was different from \(t\). After introducing
this new set of terms, the latent function now consists of a total of
\(Z=L+M+Q+R\) covariates and associated parameters.

\hypertarget{strategy-2-hierarchical-model-specification}{%
\subsection{Strategy 2: hierarchical model
specification}\label{strategy-2-hierarchical-model-specification}}

We can choose to conceptualize the event leading to the outcome as a
hierarchical process. There are a few different ways of doing this. For
example, the hierarchy could be based on individuals in traffic units.
In this case, we can rewrite the latent function as follows:

\begin{equation}
\label{eq:latent-function-with-hierarchical-traffic-unit}
y_{itk}^*=\sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \sum_{r=1}^R\delta_ro_{jvkr} + \epsilon_{itk}
\end{equation}

The coefficients of the traffic unit nest the individual attributes as
follows. For any given coefficient \(m\):

\begin{equation}
\label{eq:hierarchical-traffic-unit-coefficients}
\beta_{m}=\sum_{l=1}^L\beta_{ml}p_{itkl} 
\end{equation}

Therefore, the corresponding term in the latent function becomes
(assuming that \(p_{itk1} = 1\), i.e., it is a constant term):

\begin{equation}
\label{eq:hierarchical-traffic-unit-coefficients}
\begin{array}{rcl}\
\beta_{m}u_{tkm} &=& \big( \beta_{m1} + \beta_{m2}p_{itk2} + \cdots + \beta_{mL}p_{itkL}\big)u_{tkm}\\ 
&=& \beta_{m1}u_{tkm} + \beta_{m2}p_{itk2}u_{tkm} + \cdots + \beta_{mL}p_{itkL}u_{tkm}
\end{array}
\end{equation}

As an alternative, the nesting unit could be the interaction
person-opponent, in which case the opponent-level attributes are nested
in the following fashion:

\begin{equation}
\label{eq:latent-function-with-opponent-variables}
y_{itk}^*=\sum_{l=1}^L\alpha_lp_{itkl} + \sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \epsilon_{itk}
\end{equation}

\noindent with any person-level coefficient \(l\) that we wish to expand
defined as follows:

\begin{equation}
\label{eq:hierarchical-traffic-unit-coefficients}
\alpha_{l}=\sum_{r=1}^R\alpha_{lr}o_{jvkr}
\end{equation}

\noindent with the same conditions as before, that \(j\ne i\) is the
operator of traffic unit \(v\ne t\). The corresponding term in the
latent function is now (assuming that \(o_{jvk1}=1\), i.e., it is a
constant term):

\begin{equation}
\label{eq:hierarchical-traffic-unit-coefficients}
\begin{array}{rcl}\
\alpha_{l}p_{itkl} &=& \big(\alpha_{l1} + \alpha_{l2}o_{jvk2} + \cdots + \alpha_{lR}o_{jvkR} \big)p_{itkl}\\
&=& \alpha_{l1}p_{itkl} + \alpha_{l2}o_{jvk2}p_{itkl} + \cdots + \alpha_{lR}o_{jvkR}p_{itkl}
\end{array} 
\end{equation}

Discerning readers will identify this model specification strategy as
Casetti's expansion method (Casetti, 1972; Roorda et al., 2010). This is
a deterministic strategy for modelling contextual effects which, when
augmented with random components, becomes the well-known multi-level
modelling method (Hedeker and Gibbons, 1994, more on this in Section
\ref{sec:further-considerations}). It is worthwhile to note that
higher-order hierarchical effects are possible; for instance, individual
attributes nested within traffic units, which in turn are nested within
collisions. We do not explore higher-level hierarchies further in the
current paper.

\hypertarget{strategy-3-sample-subsetting}{%
\subsection{Strategy 3: sample
subsetting}\label{strategy-3-sample-subsetting}}

The third model specification strategy that we will consider is
subsetting the sample. This is applicable in conjunction with any of the
other strategies discussed above. In essence, we define the latent
function with restrictions as follows. Consider a continuous variable,
e.g., age of person, and imagine that we wish to concentrate the
analysis on older adults (e.g., Dissanayake and Lu, 2002). The latent
function is defined as desired (see above), however, the following
restriction is applied to the sample:

\begin{equation}
\label{eq:sampling-age}
\text{Age of individual } i \text{ in traffic unit } t \text{ in crash } k = 
\begin{cases}
\ge 65 & \text{use record } itk\\
< 65 & \text{do not use record } itk
\end{cases}
\end{equation}

Suppose instead that we are interested in crashes by or against a
specific type of traffic unit (e.g., pedestrians, Amoh-Gyimah et al.,
2017):

\begin{equation}
\label{eq:sampling-pedestrian}
\text{Road user class of individual } i \text{ in traffic unit } t \text{ in crash } k = 
\begin{cases}
\text{Pedestrian} & \text{use record } itk\\
\text{Not pedestrian} & \text{do not use record } itk
\end{cases}
\end{equation}

\noindent or:

\begin{equation}
\label{eq:sampling-pedestrian-opponent}
\text{Road user class of individual } j \text{ in traffic unit } v \text{ in crash } k = 
\begin{cases}
\text{Pedestrian} & \text{use record } jvk\\
\text{Not pedestrian} & \text{do not use record } jvk
\end{cases}
\end{equation}

More generally, for any variable \(x\) of interest:

\begin{equation}
\label{eq:sampling-general}
x_{itk} = 
\begin{cases}
\text{Condition: TRUE} & \text{use record } itk\\
\text{Condition: FALSE} & \text{do not use record } itk
\end{cases}
\end{equation}

Several conditions can be imposed to subset the sample in any way that
the analyst deems appropriate or suitable.

\hypertarget{sec:application}{%
\section{Setting for empirical assessment}\label{sec:application}}

In this section we present the setting for the empirical assessment of
the modelling strategies discussed in Section \ref{sec:methods}, namely
matters related to data and model estimation.

\hypertarget{sec:data}{%
\subsection{Data for empirical assessment}\label{sec:data}}

To assess the performance of the various modelling strategies we use
data from Canada's National Collision Database (NCDB). This database
contains all police-reported motor vehicle collisions on public roads in
Canada. Data are originally collected by provinces and territories, and
shared with the federal government, that proceeds to combine, track, and
analyze them for reporting deaths, injuries, and collisions in Canada at
the national level. The NCDB is provided by Transport Canada, the agency
of the federal government of Canada in charge of transportation policies
and programs, under the Open Government License - Canada version 2.0
{[}https://open.canada.ca/en/open-government-licence-canada{]}.

The NCDB is available from 1999. For the purpose of this paper, we use
the data corresponding to 2017, which is the most recent year available
as of this writing. Furthermore, for assessment we also use the data
corresponding to 2016. Similar to databases in other jurisdictions (see
Montella et al., 2013), the NCDB contains information pertaining to the
collision, the traffic unit(s), and the person(s) involved in a crash.
The definitions of variables in this database are shown in the Appendix
at the end of this document, in Tables
\ref{tab:ncdb-descriptives-collision},
\ref{tab:ncdb-descriptives-vehicle}, and
\ref{tab:ncdb-descriptives-person}. Notice that, compared to Table
\ref{tab:variable-categories}, environmental-factors variables and
road-factors variables are gathered under a single variable class,
namely collision-related, since they are unique for each crash.

Data are organized by person; in other words, there is one record per
individual in a collision, be they drivers, passengers, pedestrians,
etc. The only variable directly available with respect to opponents in a
collision is the number of vehicles involved (see models in Bogue et
al., 2017). Therefore, the data needs to be processed to obtain
attributes of the opposing party for each individual in a collision. The
protocol to do this is described next.

\hypertarget{data-preprocessing}{%
\subsection{Data preprocessing}\label{data-preprocessing}}

To prepare the data for analysis, in particular for Strategy 1
(opponent-related factors), we apply an initial filter, whereby we scan
the database to remove all records that are not a person (including
parked cars and other objects) or that are missing information.

After the initial filter, the database is summarized to find the number
of individual-level records that correspond to each collision (C\_CASE).
At this point, there are 32,298 collisions, involving only one (known)
individual, there are 46,483 collisions involving two parties, 19,433
collisions with three parties, 8,250 collisions involving four parties,
3,783 collisions with five parties, 1,789 collisions with six parties,
and 1,491 collisions involving seven or more parties These parties were
possibly occupants in different vehicles or were otherwise their own
traffic units (e.g., pedestrians). Accordingly, the sample includes
174,741 drivers, 61,403 passengers, 10,798 pedestrians, 5,286
bicyclists, and 6,564 motorcyclists.

The next step is to remove all collisions that involve only one party.
This still leaves numerous cases where multiple parties could have been
in a single vehicle, for instance in a collision against a stationary
object. Therefore, we proceed to use the vehicle sequence number to find
the number of vehicles involved in each collision. This reveals that
there are 20,732 collisions that involve only one vehicle but possibly
multiple individuals (i.e., driver and one or more passengers). In
addition, there are 165,520 collisions involving two vehicles (and
possibly multiple individuals). Finally, there are 40,242 collisions
with three or more vehicles.

Once we have identified the number of vehicles in each collision, we
select all cases that involve only two vehicles. The most common cases
in two-vehicle collisions are those that include drivers (40,297
collisions; this is reflective of the prevalence of single-occupant
vehicles). This is followed by cases with passengers (14,120
collisions), pedestrians (5,204 collisions), bicyclists (2,238
collisions), and motorcyclists (1,016 collisions). The distribution of
individuals per traffic unit is as follows: 80,382 individuals are coded
as being in V\_ID = 1, 76,523 individuals are coded as being in V\_ID =
2, and 7,932 individuals are coded as pedestrians. In addition, 683
individuals are coded as being in vehicles 3 through 9, despite our
earlier filter to retain only collisions with two vehicles. At this
point we select only individuals assigned to vehicles 1 or 2, as well as
pedestrians. As a result of this filter a number of cases with only one
known individual need to be removed.

\begin{table}[!h]

\caption{\label{tab:summary-opponent-information}\label{tab:summary-opponent-information}Summary of opponent interactions and outcomes by road user class}
\centering
\resizebox{\linewidth}{!}{
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{lrrrrrrrlll}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Road User Class of Opponent} & \multicolumn{3}{c}{Outcome} & \multicolumn{3}{c}{Proportion by Road User Class} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-8} \cmidrule(l{3pt}r{3pt}){9-11}
Road User Class & Driver & Pedestrian & Bicyclist & Motorcyclist & No Injury & Injury & Fatality & No Injury & Injury & Fatality\\
\midrule
\rowcolor{gray!6}  \addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{All opponents}}\\
\hspace{1em}Driver & 97582 & 7880 & 3799 & 2498 & 59180 & 52143 & 436 & 0.52953 & 0.46657 & 0.003901\\
\hspace{1em}Passenger & 35359 & 1282 & 667 & 818 & 19308 & 18667 & 151 & 0.50643 & 0.48961 & 0.003961\\
\rowcolor{gray!6}  \hspace{1em}Pedestrian & 7880 & 0 & 0 & 0 & 145 & 7507 & 228 & 0.01840 & 0.95266 & \vphantom{1} 0.028934\\
\hspace{1em}Bicyclist & 3799 & 1 & 0 & 40 & 49 & 3760 & 31 & 0.01276 & 0.97917 & 0.008073\\
\rowcolor{gray!6}  \hspace{1em}Motorcyclist & 2498 & 30 & 40 & 338 & 204 & 2598 & 104 & 0.07020 & 0.89401 & 0.035788\\
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Opponent: Driver}}\\
\hspace{1em}Driver & 97582 & 0 & 0 & 0 & 45493 & 51657 & 432 & 0.46620 & 0.52937 & 0.004427\\
\rowcolor{gray!6}  \hspace{1em}Passenger & 35359 & 0 & 0 & 0 & 16672 & 18536 & 151 & 0.47151 & 0.52422 & 0.004270\\
Pedestrian & 7880 & 0 & 0 & 0 & 145 & 7507 & 228 & 0.01840 & 0.95266 & \vphantom{1} 0.028934\\
\rowcolor{gray!6}  \hspace{1em}Bicyclist & 3799 & 0 & 0 & 0 & 43 & 3725 & 31 & 0.01132 & 0.98052 & 0.008160\\
\hspace{1em}Motorcyclist & 2498 & 0 & 0 & 0 & 98 & 2299 & 101 & 0.03923 & 0.92034 & 0.040432\\
\rowcolor{gray!6}  \addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Opponent: Pedestrian}}\\
\hspace{1em}Driver & 0 & 7880 & 0 & 0 & 7693 & 187 & 0 & 0.97627 & 0.02373 & 0.000000\\
\hspace{1em}Passenger & 0 & 1282 & 0 & 0 & 1246 & 36 & 0 & 0.97192 & 0.02808 & 0.000000\\
\rowcolor{gray!6}  \hspace{1em}Pedestrian & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - & - & \vphantom{2} -\\
\hspace{1em}Bicyclist & 0 & 1 & 0 & 0 & 0 & 1 & 0 & 0.00000 & 1.00000 & 0.000000\\
\rowcolor{gray!6}  \hspace{1em}Motorcyclist & 0 & 30 & 0 & 0 & 11 & 19 & 0 & 0.36667 & 0.63333 & 0.000000\\
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Opponent: Bicyclist}}\\
\hspace{1em}Driver & 0 & 0 & 3799 & 0 & 3706 & 93 & 0 & 0.97552 & 0.02448 & 0.000000\\
\rowcolor{gray!6}  \hspace{1em}Passenger & 0 & 0 & 667 & 0 & 649 & 18 & 0 & 0.97301 & 0.02699 & 0.000000\\
\hspace{1em}Pedestrian & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - & - & \vphantom{1} -\\
\rowcolor{gray!6}  \hspace{1em}Bicyclist & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - & - & -\\
\hspace{1em}Motorcyclist & 0 & 0 & 40 & 0 & 16 & 24 & 0 & 0.40000 & 0.60000 & 0.000000\\
\rowcolor{gray!6}  \addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Opponent: Motorcyclist}}\\
\hspace{1em}Driver & 0 & 0 & 0 & 2498 & 2288 & 206 & 4 & 0.91593 & 0.08247 & 0.001601\\
\hspace{1em}Passenger & 0 & 0 & 0 & 818 & 741 & 77 & 0 & 0.90587 & 0.09413 & 0.000000\\
\rowcolor{gray!6}  Pedestrian & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - & - & \vphantom{2} -\\
\hspace{1em}Bicyclist & 0 & 0 & 0 & 40 & 6 & 34 & 0 & 0.15000 & 0.85000 & 0.000000\\
\rowcolor{gray!6}  \hspace{1em}Motorcyclist & 0 & 0 & 0 & 338 & 79 & 256 & 3 & 0.23373 & 0.75740 & 0.008876\\
\bottomrule
\end{tabular}}
\end{table}

At this point we have a complete, workable sample of individual records
of parties in two-vehicle collisions. There are two possible cases for
the collisions, depending on the traffic units involved: 1) vehicle vs
vehicle collisions (``vehicle'' is all motorized vehicles, including
motorcycles/mopeds, as well as bicycles); and 2) vehicle vs pedestrian
collisions. To identify the opposite parties in each collision it is
convenient to classify collisions by pedestrian involvement. In this
way, we find that the database includes 16,636 collisions that are
vehicle vs pedestrian (possibly multiple pedestrians), and 147,594
collisions that involve two vehicles. After splitting the database
according to pedestrian involvement, we can now extract relevant
information about the different parties in the collision. This involves
renaming the person-level variables so that we can distinguish each
individual by their party in a given record. Notice that when working
with individuals in vehicles, only drivers are considered opposites in a
collision.

Once the personal attributes of opposite operators in a given collision
are extracted, their information is joined to the individual records by
means of the collision unique identifier. As a result of this process, a
new set of variables are now available for analysis: the age, sex, and
road user class of the opposite driver, as well as the type of the
opposite vehicle. A summary of opponent interactions and outcomes can be
found in Table \ref{tab:summary-opponent-information}. The information
there shows that the most common type of opponent for drivers are other
drivers, followed by pedestrians. The only opponents of pedestrians, on
the other hand, are drivers. Bicyclists and motorcyclists are mostly
opposed by drivers, but occasionally by other road users as well. In
terms of outcomes, we observe that virtually all fatalities occur when
the opponent is a driver, and only very rarely when the opponent is a
motorcyclist. Injuries are also more common when the opponent is a
driver, whereas ``no injury'' is a relatively more frequent outcome when
the opponent is a pedestrian or a bicyclist.

\hypertarget{sec:model-estimation}{%
\subsection{Model estimation}\label{sec:model-estimation}}

Before model estimation, the variables are prepared as follows. First,
age is scaled from years to decades. Secondly, new variables are defined
to describe the vehicle type. Three classes of vehicle types are
considered: 1) light duty vehicles (which in Canada include passenger
cars, passenger vans, light utility vehicles, and light duty pick up
trucks); 2) light trucks (all other vehicles \(\le\) 4536 kg in gross
vehicle weight rating); and heavy vehicles (all other vehicles \(\ge\)
4536 kg in gross vehicle weight rating). Furthermore, this typology of
vehicle is combined with the road user class of the individual to
distinguish between drivers and passengers of light duty vehicles, light
trucks, and heavy vehicles, in addition to pedestrians, bicyclists, and
motorcyclists. This is done for both the individual and the opponent.
Variable interactions are calculated to produce hierarchical variables.
For example, for a hierarchical definition of traffic unit-level
variables, age (and the square of age to account for possible
non-monotonic effects) are interacted with gender, road user class, and
vehicle type. For hierarchical opponent variables, age (and the square
of age) are interacted with the age of opponent (and the corresponding
square). The variables thus obtained are shown in Table
\ref{tab:model-specification-summary}. As seen in the table, Models 1
and 2 are single-level models, and the difference between them is that
Model 2 includes opponent variables. Models 3 and 4, in contrast, are
hierarchical models. Model 3 considers the hierarchy on the basis of the
traffic unit, while Model 4 considers the hierarchy on the basis of the
opponent.

Models 1 through 4 are estimated using the full sample. As discussed
above, a related modelling strategy is to subset the sample (e.g., Islam
et al., 2014; Lee and Li, 2014; Torrao et al., 2014; Wu et al., 2014).
In this case we subset by a combination of traffic unit type of the
individual (i.e., light duty vehicle, light truck, heavy vehicle,
pedestrian, bicyclist, and motorcyclist) and vehicle type of the
opponent (i.e., light duty vehicle, light truck, heavy vehicle). This
leads to an ensemble of eighteen models to be estimated using subsets of
data (see Table \ref{tab:model-specification-summary}). By subsetting
the sample, at least \emph{some} opponent effects are incorporated
implicitly. Models 1 and 2 are re-estimated using this strategy,
dropping variables as necessary whenever they become irrelevant (for
instance, after filtering for pedestrians, no other traffic unit types
are present in the subset of data). In addition to variables that are no
longer relevant in some data subsets, it is important to note that when
using some data subsets a few variables had to be occasionally dropped
to avoid convergence issues. This tended to happen particularly with
smaller subsets where some particular combination of attributes was rare
as a result of subsampling (e.g., in 2017 there were few or no
collisions that involved a motorcyclist and a heavy vehicle in a bridge,
or overpass, or viaduct). The process of estimation carefully paid
attention to convergence issues to ensure the validity of the models
reported here.

\begin{table}[!h]

\caption{\label{tab:model-specification-summary}\label{tab:model-specification-summary}Summary of variables and model specification}
\centering
\resizebox{\linewidth}{!}{
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{ll>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{6em}}
\toprule
Variable & Notes & Model 1
Single-level
/No opponent & Model 2
Single-level
/Opponent attributes & Model 3
Hierarchical: Traffic unit & Model 4
Hierarchical: Opponent attributes\\
\midrule
\rowcolor{gray!6}  \addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Individual-level variables}}\\
\hspace{1em}Age & In decades & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\hspace{1em}Age Squared &  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\rowcolor{gray!6}  \hspace{1em}Sex & Reference: Female & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\hspace{1em}Use of Safety Devices & 7 levels; Reference: No Safety Device & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\rowcolor{gray!6}  \addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Traffic unit-level variables}}\\
\hspace{1em}Passenger & Reference: Driver & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
\hspace{1em}Pedestrian & Reference: Driver & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
\rowcolor{gray!6}  \hspace{1em}Bicyclist & Reference: Driver & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
\hspace{1em}Motorcyclist & Reference: Driver & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
\rowcolor{gray!6}  \hspace{1em}Light Truck & Reference: Light Duty Vehicle & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
\hspace{1em}Heavy Vehicle & Reference: Light Duty Vehicle & $\checkmark$ & $\checkmark$ &  & $\checkmark$\\
\rowcolor{gray!6}  \addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Opponent variables}}\\
\hspace{1em}Age of Opponent & In decades &  & $\checkmark$ & $\checkmark$ & \\
\hspace{1em}Age of Opponent Squared &  &  & $\checkmark$ & $\checkmark$ & \\
\rowcolor{gray!6}  \hspace{1em}Sex of Opponent & Reference: Female &  & $\checkmark$ & $\checkmark$ & \\
\hspace{1em}Opponent: Light Duty Vehicle & Reference: Pedestrian/Bicyclist/Motorcyclist &  & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\rowcolor{gray!6}  \hspace{1em}Opponent: Light Truck & Reference: Pedestrian/Bicyclist/Motorcyclist &  & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\hspace{1em}Opponent: Heavy Vehicle & Reference: Pedestrian/Bicyclist/Motorcyclist &  & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\rowcolor{gray!6}  \addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Hierarchical traffic unit variables}}\\
\hspace{1em}Light Truck Driver:Age &  &  &  & $\checkmark$ & \\
\hspace{1em}Light Truck Driver:Age Squared &  &  &  & $\checkmark$ & \\
\rowcolor{gray!6}  \hspace{1em}Heavy Vehicle Driver:Age &  &  &  & $\checkmark$ & \\
\hspace{1em}Heavy Vehicle Driver:Age Squared &  &  &  & $\checkmark$ & \\
\rowcolor{gray!6}  \hspace{1em}Light Truck Passenger:Age &  &  &  & $\checkmark$ & \\
\hspace{1em}Light Truck Passenger:Age Squared: &  &  &  & $\checkmark$ & \\
\rowcolor{gray!6}  \hspace{1em}Heavy Vehicle Passenger:Age &  &  &  & $\checkmark$ & \\
\hspace{1em}Heavy Vehicle Passenger:Age Squared &  &  &  & $\checkmark$ & \\
\rowcolor{gray!6}  \hspace{1em}Pedestrian:Age &  &  &  & $\checkmark$ & \\
\hspace{1em}Pedestrian:Age Squared &  &  &  & $\checkmark$ & \\
\rowcolor{gray!6}  \hspace{1em}Bicyclist:Age &  &  &  & $\checkmark$ & \\
\hspace{1em}Bicyclist:Age Squared &  &  &  & $\checkmark$ & \\
\rowcolor{gray!6}  \hspace{1em}Motorcyclist:Age &  &  &  & $\checkmark$ & \\
\hspace{1em}Motorcyclist:Age Squared &  &  &  & $\checkmark$ & \\
\rowcolor{gray!6}  \addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Hierarchical opponent variables}}\\
\hspace{1em}Age:Age of Opponent &  &  &  &  & $\checkmark$\\
\hspace{1em}Age:Age of Female Opponent &  &  &  &  & $\checkmark$\\
\rowcolor{gray!6}  \hspace{1em}Age:Age of Male Opponent Squared &  &  &  &  & $\checkmark$\\
\hspace{1em}Age:Age of Female Opponent Squared &  &  &  &  & $\checkmark$\\
\rowcolor{gray!6}  \hspace{1em}Age Squared:Age of Male Opponent &  &  &  &  & $\checkmark$\\
\hspace{1em}Age Squared:Age of Female Opponent &  &  &  &  & $\checkmark$\\
\rowcolor{gray!6}  \addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Collision-level variables}}\\
\hspace{1em}Crash Configuration & 19 levels; Reference: Hit a moving object & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\hspace{1em}Road Configuration & 12 levels; Reference: Non-intersection & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\rowcolor{gray!6}  \hspace{1em}Weather & 9 levels; Reference: Clear and sunny & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\hspace{1em}Surface & 11 levels; Reference: Dry & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\rowcolor{gray!6}  \hspace{1em}Road Alignment & 8 levels; Reference: Straight and level & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\hspace{1em}Traffic Controls & 19 levels; Reference: Operational traffic signals & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\rowcolor{gray!6}  \hspace{1em}Month & 12 levels; Reference: January & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{sec:assessment}{%
\section{Model assessment}\label{sec:assessment}}

In this section we report an in-depth examination of the performance of
the models. We begin by inspecting the statistical goodness of fit of
the models by means of Akaike's Information Criterion (\(AIC\)). Next,
we use the models to conduct in-sample predictions (i.e.,
\emph{nowcasting}), using the same sample that was used to estimate the
models, and out-of-sample predictions, using the data set corresponding
to the year 2016 (i.e., \emph{backcasting}). Predictions are commonly
evaluated in two different ways in the literature. Some researchers
analyze the outcome shares based on the predicted probabilities (e.g.,
Bogue et al., 2017; Yasmin and Eluru, 2013). This is a form of aggregate
forecasting. Other researchers, in contrast, evaluate the classes of
outcomes based on the individual-level predictions (e.g., Tang et al.,
2019; Torrao et al., 2014). This is a form of disaggregate forecasting.

\hypertarget{sec:goodness-of-fit}{%
\subsection{Goodness of fit of models}\label{sec:goodness-of-fit}}

We begin our empirical assessment by examining the results of estimating
the models described above. Tables \ref{tab:model-full-sample-summary}
and \ref{tab:model-subsample-summary} present some key summary
statistics of the estimated models. Of interest is the goodness of fit
of the models, which in the case is measured with Akaike's Information
Criterion (\(AIC\)). This criterion is calculated as follows:

\begin{equation}
\label{eq:aic}
AIC = 2Z - 2\ln{\hat{L}}
\end{equation}

\noindent where \(Z\) is the number of coefficients estimated by the
model, and \(\hat{L}\) the maximized likelihood of the model. Since
\(AIC\) penalizes the model fit by means of the number of coefficients,
this criterion gives preference to more parsimonious models. The
objective is to minimize the \(AIC\), and therefore smaller values of
this criterion represent better model fits. Model comparison can be
conducted using the relative likelihood. Suppose that we have two
models, say Model \(a\) and Model \(b\), with \(AIC_{a} \le AIC_{b}\).
The relative likelihood is calculated as:

\begin{equation}
\label{eq:relative-likelihood}
e^{(AIC_a - AIC_b)/2}
\end{equation}

The relative likelihood is interpreted as the probability that Model
\(b\) minimizes the information loss as well as Model \(a\).

Turning our attention to the models estimated using the full sample
(Table \ref{tab:model-full-sample-summary}), it is possible to see that,
compared to the base (single-level) model without opponent variables
(Model 1), there are large and significant improvements in goodness of
fit to be gained by introducing opponent effects. However, the gains are
not as large when hierarchical specifications are used, even when the
number of additional coefficients that need to be estimated is not
substantially larger (recall that the penalty per coefficient in \(AIC\)
is 2). The best model according to this measure of goodness of fit is
Model 2 (single-level with opponent effects), followed by Model 4
(hierarchical opponent variables), Model 3 (hierarchical traffic unit
variables with opponent effects), and finally Model 1 (single-level
without opponent effects).

It is important to note that the likelihood function of a model, and
therefore the value of its \(AIC\), both depend on the size of the
sample, which is why \(AIC\) is not comparable across models estimated
with different sample sizes. For this reason, the full sample models
cannot be compared directly to the models estimated with subsets of
data. The models in the ensembles, however, can be compared to each
other (Table \ref{tab:model-subsample-summary}). As seen in the table,
introducing opponent variables leads to a better fit in the case of
most, but not all models. The simplest model (single-level without
opponent effects) is clearly the best fitting candidate in the case of
bicycle vs light truck collisions, bicycle vs heavy vehicle collisions,
motorcyclist vs light duty vehicle collisions, and motorcyclist vs heavy
vehicle collisions. Model 1 is a statistical toss for best performance
with two competing models in the case of pedestrian vs heavy vehicle
collisions. The relative likelihood of Model 1 compared to Models 2 and
3 in this case is 0.56, which means that these two models are 0.56 times
as probable as Model 1 to minimize the information loss.

Model 2 is the best fit in the case of light truck vs heavy vehicle
collisions. This model is also tied for best fit with Model 2 in the
case of pedestrian vs light duty vehicle and pedestrian vs light truck
collisions, and is a statistical toss with Model 4 in the case of heavy
vehicle vs heavy vehicle collisions (relative likelihood is 0.592).
Model 3 is the best fit in the case of light duty vehicle vs light duty
vehicle collisions and heavy vehicle vs light duty vehicle. Model 4 is
the best fit in the case of light duty vehicle vs light truck
collisions, light duty vehicle vs heavy vehicle collisions, light truck
vs light duty vehicle collisions, heavy vehicle vs light truck
collisions, and motorcycle vs light truck collisions. This model is a
statistical toss with Model 2 in the case of light truck vs light truck
collisions, with a relative likelihood of 0.791.

These results give some preliminary ideas about the relative performance
of the different modelling strategies. In the next subsection we delve
more deeply into this question by examining the predictive performance
of the various modelling strategies. The results up to this point
indicate that different model specification strategies might work best
when combined with subsampling strategies. For space reasons, from this
point onwards, we will consider the ensembles of models for predictions
and will not compare individual models within the ensembles; this we
suggest is a matter for future research.

\begin{table}

\caption{\label{tab:model-full-sample-summary}\label{tab:model-full-sample-summary}Summary of model estimation results: Full sample models}
\centering
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{lccc}
\toprule
Model & \makecell[l]{Number of\\observations} & \makecell[l]{Number of\\coefficients} & AIC\\
\midrule
\rowcolor{gray!6}  Model 1. Single-level/No opponent & 164,511 & 102 & 195,215\\
Model 2. Single-level/Opponent attributes & 164,511 & 108 & 178,943\\
\rowcolor{gray!6}  Model 3. Hierarchical: Traffic unit & 164,511 & 118 & 181,333\\
Model 4. Hierarchical: Opponent) & 164,511 & 111 & 179,018\\
\bottomrule
\end{tabular}
\end{table}

\begin{landscape}\begin{table}

\caption{\label{tab:model-subsample-summary}\label{tab:model-summary}Summary of model estimation results: Data Subset Models}
\centering
\resizebox{\linewidth}{!}{
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{lccccccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{2}{c}{\makecell[c]{Model 1\\Single-level/No opponent}} & \multicolumn{2}{c}{\makecell[c]{Model 2\\Single-level/Opponent attributes}} & \multicolumn{2}{c}{\makecell[c]{Model 3\\Hierarchical: Traffic unit}} & \multicolumn{2}{c}{\makecell[c]{Model 4\\Hierarchical: Opponent}} \\
\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8} \cmidrule(l{3pt}r{3pt}){9-10}
Model & \makecell[l]{Number of\\observations} & \makecell[l]{Number of\\coefficients} & AIC & \makecell[l]{Number of\\coefficients} & AIC & \makecell[l]{Number of\\coefficients} & AIC & \makecell[l]{Number of\\coefficients} & AIC\\
\midrule
\rowcolor{gray!6}  Light duty vehicle vs light duty vehicle & 114,841 & 94 & 145,390 & 97 & 143,903 & 100 & 143,896 & 100 & 144,004\\
Light duty vehicle vs light truck & 3,237 & 79 & 3,943 & 82 & 3,927 & 85 & 3,937 & 85 & 3,922\\
\rowcolor{gray!6}  Light duty vehicle vs heavy vehicle & 5,013 & 88 & 5,895 & 91 & 5,878 & 94 & 5,888 & 94 & 5,864\\
Light truck vs light duty vehicle & 3,121 & 79 & 3,885 & 82 & 3,877 & 85 & 3,881 & 85 & 3,875\\
\rowcolor{gray!6}  Light truck vs light truck & 809 & 67 & 1,170 & 70 & 1,156 & 73 & 1,162 & 73 & 1,155\\
\addlinespace
Light truck vs heavy vehicle & 198 & 64 & 288 & 67 & 281 & 70 & 287 & 70 & 286\\
\rowcolor{gray!6}  Heavy vehicle vs light duty vehicle & 4,726 & 79 & 4,326 & 84 & 4,283 & 86 & 4,268 & 87 & 4,287\\
Heavy vehicle vs light truck & 180 & 64 & 225 & 65 & 205 & 67 & 207 & 66 & 187\\
\rowcolor{gray!6}  Heavy vehicle vs heavy vehicle & 779 & 74 & 1,147 & 77 & 1,136 & 80 & 1,141 & 80 & 1,137\\
Pedestrian vs light duty vehicle & 7,176 & 88 & 2,826 & 91 & 2,821 & 91 & 2,821 & 93 & 2,827\\
\addlinespace
\rowcolor{gray!6}  Pedestrian vs light truck & 328 & 62 & 202 & 65 & 200 & 65 & 200 & 68 & 206\\
Pedestrian vs heavy vehicle & 376 & 64 & 409 & 67 & 410 & 67 & 410 & 70 & 417\\
\rowcolor{gray!6}  Bicyclist vs light duty vehicle & 3,521 & 80 & 654 & 83 & 659 & 83 & 659 & 85 & 657\\
Bicyclist vs light truck & 116 & 42 & 84 & 57 & 114 & 57 & 114 & 54 & 108\\
\rowcolor{gray!6}  Bicyclist vs heavy vehicle & - & - & - & - & - & - & - & - & -\\
\addlinespace
Motorcyclist vs light duty vehicle & 2,298 & 78 & 1,367 & 81 & 1,373 & 81 & 1,373 & 84 & 1,373\\
\rowcolor{gray!6}  Motorcyclist vs light truck & 127 & 56 & 153 & 59 & 153 & 59 & 153 & 47 & 94\\
Motorcyclist vs heavy vehicle & 62 & 43 & 88 & 45 & 90 & 46 & 92 & 51 & 102\\
\bottomrule
\multicolumn{10}{l}{\textit{Note: }}\\
\multicolumn{10}{l}{There are zero cases of Bicyclist vs heavy vehicle in the sample}\\
\end{tabular}}
\end{table}
\end{landscape}

\hypertarget{sec:outcomes-shares}{%
\subsection{Outcome shares based on predicted
probabilities}\label{sec:outcomes-shares}}

In this, and the following section, \emph{backcasting} refers to the
prediction of probabilities and classes of outcomes using the 2016 data
set. When conducting backcasting, the data set is preprocessed in
identical manner as the 2017 data set. In addition, the variables used
in backcasting match exactly those in the models. This means that some
variables were dropped when they were present in the 2016 data set but
not in the models. This tended to happen in the case of relatively rare
outcomes (e.g., in 2016, there was at least one collision between a
heavy vehicle and a light duty vehicle in a school crossing zone; no
such event was observed in 2017).

The shares of each outcome are calculated as the sum of the estimated
probabilities for each observation:

\begin{equation}
\label{eq:predicted-shares}
\begin{array}{rcl}
\hat{S}_{\text{PDO}} &=& \sum_{itk}\hat{P}(y_{itk}=\text{PDO})\\
\hat{S}_{\text{injury}} &=& \sum_{itk}\hat{P}(y_{itk}=\text{injury})\\
\hat{S}_{\text{fatality}} &=& \sum_{itk}\hat{P}(y_{itk}=\text{fatality})
\end{array}
\end{equation}

\noindent where \(\hat{P}(y_{itk}=h_w)\) is the estimated probability of
outcome \(h_w\) for individual \(i\) in traffic unit \(t\) and crash
\(k\). The estimated share of outcome \(h\) is \(\hat{S}_{h_w}\).

The estimated shares can be used to assess the ability of the model to
forecast for the population the total number of cases of each outcome. A
summary statistic useful to evaluate the performance is the Average
Percentage Error, or \(APE\) (see Bogue et al., 2017, p. 31), which is
calculated for each outcome as follows:

\begin{equation}
\label{eq:APE}
APE_{h_w} = \Bigg|\frac{\hat{S}_{h_w} - S_{h_w}}{S_{h_w}}\Bigg|\times 100
\end{equation}

The Weighted Average Percentage Error (\(WAPE\)) aggregates the \(APE\)
as follows:

\begin{equation}
\label{eq:WAPE}
WAPE = \frac{\sum_w^WAPE_{h_w}\times S_{h_w}}{\sum_w^WS_{h_w}}
\end{equation}

The results of this exercise are reported in Table
\ref{tab:ape-results}. Of the four full-sample models (Models 1-4), the
\(APE\) of Model 2 is lowest in the nowcasting exercise for every
outcome, with the exception of Fatality, where Model 4 produces a
considerably lower \(APE\). When the results are aggregated by means of
the \(WAPE\), Model 2 gives marginally better results than Model 4. It
is interesting to see that the four ensemble models have lower \(APE\)
values across the board in the nowcasting exercise, and much better
\(WAPE\) than the full sample models. However, once we turn to the
results of the backcasting exercise, these results do not hold, and it
is possible to see that the Average Percentage Errors of the ensemble
worsen considerably, particularly in the case of Fatality. The Weighted
Average Prediction Error of the ensemble models in the backcasting
exercise is also worse than for any of the full sample models. Excellent
in-sample predictions but mediocre out-of-sample predictions are often
evidence of overfitting, as in the case of the ensemble models here.

In terms of backcasting, full sample Model 1 is marginally better than
full sample Models 2 and 3, and better than full sample Model 4. The
reason for this is the lower \(APE\) of Model 1 when predicting Injury,
the most frequent outcome. However, the performance of Model 1 with
respect to Fatality (the least frequent outcome) is the worst of all
models. Whereas Model 4 has the best performance predicting Fatality,
its performance with respect to other classes of outcomes is less
impressive. Model 3 does better than Model 2 with respect to Injury, but
performs relatively poorly when backcasting Fatality. Overall, Model 2
appears to be the most balanced, with good in-sample performance and
competitive out-of-sample performance that is also balanced with respect
to the various classes of outcomes.

\begin{table}

\caption{\label{tab:table-ape-results}\label{tab:ape-results}Predicted shares and average prediction errors (APE) by model (percentages)}
\centering
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{lrrrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{3}{c}{No Injury} & \multicolumn{3}{c}{Injury} & \multicolumn{3}{c}{Fatality} & \multicolumn{1}{c}{} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10}
Model & Observed & Predicted & APE & Observed & Predicted & APE & Observed & Predicted & APE & WAPE\\
\rowcolor{gray!15}
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{In-sample (nowcasting using 2017 data set, i.e., estimation data set)}}\\
\hspace{1em}Model 1 & 78886 & 79029.00 & 0.18 & 84675 & 84533.74 & 0.17 & 950 & 948.26 & 0.18 & 0.17\\
\rowcolor{gray!15}
\hspace{1em}Model 2 & 78886 & 78928.98 & 0.05 & 84675 & 84641.94 & 0.04 & 950 & 940.08 & 1.04 & 0.05\\
\hspace{1em}Model 3 & 78886 & 79027.29 & 0.18 & 84675 & 84512.50 & 0.19 & 950 & 971.21 & 2.23 & 0.20\\
\rowcolor{gray!15}
\hspace{1em}Model 4 & 78886 & 78939.18 & 0.07 & 84675 & 84622.54 & 0.06 & 950 & 949.28 & 0.08 & 0.06\\
\hspace{1em}Model 1 Ensemble & 62413 & 62402.78 & 0.02 & 83564 & 83573.58 & 0.01 & 931 & 931.64 & 0.07 & 0.01\\
\rowcolor{gray!15}
\hspace{1em}Model 2 Ensemble & 62417 & 62407.00 & 0.02 & 83595 & 83604.14 & 0.01 & 931 & 931.86 & 0.09 & 0.01\\
\rowcolor{gray!15}
\hspace{1em}Model 3 Ensemble & 62411 & 62401.23 & 0.02 & 83596 & 83604.71 & 0.01 & 933 & 934.06 & 0.11 & 0.01\\
\hspace{1em}Model 4 Ensemble & 62405 & 62395.28 & 0.02 & 83578 & 83586.75 & 0.01 & 932 & 932.97 & 0.10 & 0.01\\
\rowcolor{gray!15}
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Out-of-sample (backcasting using 2016 data set)}}\\
\hspace{1em}Model 1 & 96860 & 96364.67 & 0.51 & 101605 & 102002.59 & 0.39 & 1109 & 1206.74 & 8.81 & 0.50\\
\rowcolor{gray!15}
\hspace{1em}Model 2 & 96860 & 96361.41 & 0.51 & 101605 & 102112.08 & 0.50 & 1109 & 1100.51 & 0.77 & 0.51\\
\hspace{1em}Model 3 & 96860 & 96354.01 & 0.52 & 101605 & 102086.18 & 0.47 & 1109 & 1133.82 & 2.24 & 0.51\\
\rowcolor{gray!15}
\hspace{1em}Model 4 & 96860 & 96325.85 & 0.55 & 101605 & 102136.72 & 0.52 & 1109 & 1111.43 & 0.22 & 0.54\\
\hspace{1em}Model 1 Ensemble & 77457 & 76822.49 & 0.82 & 100013 & 100580.60 & 0.57 & 1072 & 1138.91 & 6.24 & 0.71\\
\rowcolor{gray!15}
\hspace{1em}Model 2 Ensemble & 77459 & 76799.11 & 0.85 & 100049 & 100630.48 & 0.58 & 1071 & 1149.41 & 7.32 & 0.74\\
\hspace{1em}Model 3 Ensemble & 77459 & 76786.76 & 0.87 & 100050 & 100644.29 & 0.59 & 1072 & 1149.95 & 7.27 & 0.75\\
\rowcolor{gray!15}
Model 4 Ensemble & 77461 & 76766.08 & 0.90 & 100029 & 100630.21 & 0.60 & 1070 & 1163.71 & 8.76 & 0.78\\
\bottomrule
\multicolumn{11}{l}{\textit{Note: }}\\
\multicolumn{11}{l}{Model 1. Single-level/No opponent}\\
\multicolumn{11}{l}{Model 2. Single-level/Opponent attributes}\\
\multicolumn{11}{l}{Model 3. Hierarchical: Traffic unit}\\
\multicolumn{11}{l}{Model 4. Hierarchical: Opponent}\\
\end{tabular}
\end{table}

\hypertarget{sec:outcomes-classes}{%
\subsection{Outcome frequency based on predicted
classes}\label{sec:outcomes-classes}}

\(APE\) and \(WAPE\) are summary measures of the performance of models
at the aggregated level. Aggregate-level predictions (i.e., shares of
outcomes) are of interest from a population health perspective. In other
cases, an analyst might be interested in the predicted outcomes at the
individual level. In this section we examine the frequency of outcomes
based on predicted classes, using the same two settings as above:
nowcasting and backcasting.

The individual-level outcomes are examined using an array of
verification statistics. Verification statistics are widely used in the
evaluation of predictive approaches were the outcomes are categorical,
and are often based on the analysis of \emph{confusion matrices} (e.g.,
Provost and Kohavi, 1998; Beguería, 2006). Confusion matrices are
cross-tabulations of \emph{observed} and \emph{predicted} classes. In a
two-by-two confusion matrix there are four possible combinations of
observed to predicted classes: hits, misses, false alarms, and correct
non-events, as shown in Table \ref{tab:confusion-matrix}. When the
outcome has more than two classes, the confusion matrix is converted to
a two-by-two table to calculate verification statistics.

\begin{table}[!h]

\caption{\label{tab:example-confusion-matrix}\label{tab:confusion-matrix}Example of a two-by-two confusion matrix}
\centering
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{lccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Observed} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-3}
Predicted & Yes & No & Marginal Total\\
\midrule
\rowcolor{gray!6}  Yes & Hit & False Alarm & Predicted Yes\\
No & Miss & Correct Non-event & Predicted No\\
\rowcolor{gray!6}  Marginal Total & Observed Yes & Observed No & \\
\bottomrule
\end{tabular}
\end{table}

The statistics used in our assessment are summarized in Table
\ref{tab:verification-statistics}, including brief descriptions of their
interpretation. The statistics evaluate different aspects of the
performance of a model. Some are concerned with the ability of the model
to be right, others are concerned with the ability of the model to match
the observed outcomes, and yet others measure the ability of the model
to not be wrong. These verification statistics are discussed briefly
next.

Percent correct and percent correct by class \(PC\) and \(PC_c\) are
relatively simple statistics, and are calculated as the proportion of
correct predictions (i.e., hits and correct non-events) relative to the
number of events, for the whole table \(PC\) or for one class only
\(PC_c\).

Bias (\(B\)) measures for each outcome class the proportion of total
predictions by class (e.g., hits as well as false alarms) relative to
the total number of cases observed for that class. For this reason, it
is possible for predictions to have low bias (values closer to 1) but
still do poorly in terms of hits.

Critical Success Index (\(CSI\)) evaluates forecasting skill while
assuming that the number of correct non-events is inconsequential for
demonstrating skill. Accordingly, the statistic is calculated as the
proportion of hits relative to the sum of hits plus false alarms plus
misses.

Probability of False Detection (\(F\)) is the proportion of false alarms
relative to the total number of times that the event is not observed.
This statistic measures the frequency with which the model incorrectly
predicts an event, but not when it incorrectly misses it. The
Probability of Detection (\(POD\)), in contrast, measures the frequency
with which the model correctly predicts a class, relative to the number
of cases of that class.

The False Alarm Ratio (\(FAR\)) is the fraction of predictions by class
that were false alarms evaluate a different way in which a model can
make equivocal predictions. In this case, lower scores are better.

The last three verification statistics that we consider are skill scores
that simultaneously consider different aspects of prediction, and are
therefore overall indicators of prediction skill. Heidke's Skill Score
(\(HSS\)) is the fraction of correct predictions above those that could
be attributed to chance. Peirce's Skill Score (\(PSS\)) combines the
Probability of Detection (\(POD\)) of a model and its Probability of
False Detection (\(F\)) to measure the skill of a model to discriminate
the classes of outcomes. Lastly, Gerrity Score (\(GS\)) is a measure of
the model's skill predicting the correct classes that tends to reward
correct forecasts of the least frequent class.

We discuss the results of calculating this battery of verification
statistics, first for the nowcasting case (Table
\ref{tab:nowcast-outcomes-assessment}) and subsequently for the
backcasting case (Table \ref{tab:backcast-outcomes-assessment}).

\begin{landscape}\begin{table}

\caption{\label{tab:table-verification-statistics}\label{tab:verification-statistics}Verification statistics}
\centering
\resizebox{\linewidth}{!}{
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{30em}>{\raggedright\arraybackslash}p{30em}}
\toprule
Statistic & Description & Notes\\
\midrule
\rowcolor{gray!6}  Percent Correct ($PC$) & Total hits and correct non-events divided by number of cases & Strongly influenced by most common category\\
Percent Correct by Class ($PC_c$) & Same as Percent Correct but by category & Strongly influenced by most common category\\
\rowcolor{gray!6}  Bias ($B$) & Total predicted by category, divided by total observed by category & $B>1$: class is overpredicted; $B<1$: class is underpredicted\\
Critical Success Index ($CSI$) & Total hits divided by total hits + false alarms + misses & $CSI = 1$: perfect score; $CSI = 0$: no skill\\
\rowcolor{gray!6}  Probability of False Detection ($F$) & Proportion of no events forecast as yes; sensitive to false alarms but ignores misses & $F = 0$: perfect score\\
\addlinespace
Probability of Detection ($POD$) & Total hits divided by total observed by class & $POD = 1$: perfect score\\
\rowcolor{gray!6}  False Alarm Ratio ($FAR$) & Total false alarms divided by total forecast yes by class; measures fraction of predicted yes that did not occur & $FAR = 0$: perfect score\\
Heidke Skill Score ($HSS$) & Fraction of correct predictions after removing predictions attributable to chance; measures fractional improvement over random; tends to reward conservative forecasts & $HSS = 1$: perfect score; $HSS = 0$: no skill; $HSS < 0$: random is better\\
\rowcolor{gray!6}  Peirce Skill Score ($PSS$) & Combines $POD$ and $F$; measures ability to separate yes events from no events; tends to reward conservative forecasts & $PSS = 1$: perfect score; $PSS = 0$: no skill\\
Gerrity Score ($GS$) & Measures accuracy of predicting the correct category, relative to random; tends to reward correct forecasts of less likely category & $GS = 1$: perfect score; $GS = 0$: no skill\\
\bottomrule
\end{tabular}}
\end{table}
\end{landscape}

\hypertarget{nowcasting-verification-statistics}{%
\subsubsection{Nowcasting: verification
statistics}\label{nowcasting-verification-statistics}}

At first glance, the results of the verification statistics (Table
\ref{tab:nowcast-outcomes-assessment}) make it clear that no model under
comparison is consistently a top performer from every aspect of
prediction. Recalling Box's aphorism, all models are wrong but some are
useful - in this case it just so happens that some models are more wrong
than others in subtly different ways. That said, it is noticeable that
the worst scores across the board tend to accrue to Model 1 in its full
sample and ensemble versions. On the other hand, Model 2 (full sample)
concentrates most of the best scores and second best scores of all the
models, but also some of the worst scores for Fatality. Model 4, in
contrast, has most of the second best scores and a few top scores, but
not a single worst score.

Of all the models, Model 2 (full sample) performs best in terms of
Percent Correct, followed by Model 4 (full sample). The worst performer
from this perspective is Model 1 (full sample), with a \(PC\) score
several percentage points below the top models. The second score is
Percent Correct by Class (\(PC_c\)). This score is calculated
individually for each outcome class. Model 2 (full sample), has the best
performance for outcomes No Injury and Injury, and the second best score
for Fatality. Model 4 (full sample) has the best score for Fatality, and
is second best for No Injury and Injury. Model 1 (full sample) has worst
scores for No Injury and Injury whereas its ensemble version has the
worst score for Fatality. It is important to note that \(PC\) and
\(PC_c\) are heavily influenced by the most common category, something
that can be particularly appreciated in the scores for Fatality. The
scores for this class of outcome are generally high, despite the fact
that the number of hits are relatively low; the high values of \(PC_c\)
in this case are due to the high occurrence of correct non-events
elsewhere in the table.

The models with the best performance in terms of \(B\) are Model 1 (full
sample) for No Injury and Injury, and Model 3 (ensemble) for Fatality.
Model 4 (full sample) is the second best performer for No Injury and
Injury, and Model 4 (ensemble) is second best performer for Fatality.
Model 1 (ensemble) has the worst bias for No Injury and Injury, whereas
Model 2 (full sample) has the worst bias for Fatality.

No model performs uniformly best from the perspective of Critical
Success Index (\(CSI\)). Model 2 (full sample) has the best \(CSI\) for
No Injury, Model 2 (ensemble) has the best score for Injury, and Model 4
(ensemble) the best score for Fatality. On the other hand, Model 1
(ensemble) has the worst score for No Injury, Model 1 (full sample) the
worst score for Injury, and Model 2 the worst score for Fatality. These
scores indicates that the models are not particularly skilled at
predicting the corresponding classes correctly, given the frequency with
which they give false alarms or miss the class.

The lowest probability of false detection \(F\) in the case of No Injury
is 19.17\% for Model 2 (ensemble), with every other model having values
lower than 21\%, with the exception of Model 1 (full sample) that has a
score of 26.46\%. With respect to Injury, the lowest probabilities range
35.8\% and 35.29\% and 35.35\% for Models 4 (full sample) and 2 (full
sample). In contrast, the highest probability of false detection for
Injury is 45.12\% for Model 1 (ensemble). The scores for \(F\) for
Fatality are all extremely low as a consequence of the very low
frequency of this class of outcome in the sample.

As with some other verification statistics, no model is consistently a
best performer in terms of \(POD\). Model 4 (full sample) has the
highest probability of detection for No Injury (65.38\%), followed by
Model 2 (65.32\%), whereas the worst probability of detection is by
Model 1 (ensemble) with a score of 55.54\%. In terms of Injury, all
models have \(POD\) higher than 79\%, and the highest score is 80.67\%
for Model 2 (ensemble). The exception is Model 1 (full sample), which
has a considerably lower \(POD\) of Injury with a score of 73.36\%.
Lastly, in terms of Fatalities, all models have very low probabilities
of detection, ranging from a high of 4.94\% in the case of Model 4
(ensemble) to a worst score of 0.11\% in the case of Model 2 (full
sample).

Model 2 (full sample) has the best \(FAR\) statistic for No Injury, as
only 25.05\% of predictions for this class are false alarms. The next
best score is by Model 4 (full sample), with only 25.23\% of No Injury
predictions being false alarms. The worst performance in this class is
by Model 1 (ensemble), which produces almost a third of false alarms in
its predictions of No Injury. In the case of Injury, the False Alarm
Ratio ranges from a low of 29.15\% by Model (ensemble), with every other
model having scores lower than 30\% except Model 1 (full sample), that
gives almost 32\% of false alarms. In terms of Fatality, the lowest
\(FAR\) is also for Model 4 (ensemble) with only 17.86\% of false
alarms, whereas the worst performance is by Model 2 (full sample), which
produces over 95\% of false alarms.

The skill scores help to remove some of the ambiguity regarding the
overall performance of a model. In this way, we know that Model 2 (full
sample) does not do particularly well with the class Fatality - however,
of all models, it tends to have the best overall performance. Its
\(HSS\), for example, suggests that it achieves 44.74\% of correct
predictions after removing correct predictions attributable to chance.
In contrast, the lowest score is for Model 1 (ensemble), which only
achieves 36.26\% correct predictions after removing those attributable
to chance. Model 2 (full sample) also has the highest \(PSS\) and the
second highest \(GS\). Model 4 (full sample) has the highest \(GS\) and
the second highest \(HSS\) and \(PSS\).

\begin{table}

\caption{\label{tab:nowcast-outcomes-assessment}\label{tab:nowcast-outcomes-assessment}Assessment of in-sample outcomes (nowcasting using 2017 data set, i.e., estimation data set)}
\centering
\resizebox{\linewidth}{!}{
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{lrrrllllllllll}
\toprule
\multicolumn{1}{c}{Predicted} & \multicolumn{3}{c}{Observed Outcome} & \multicolumn{10}{c}{Verification Statistics} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-14}
Outcome & No Injury & Injury & Fatality & \makecell[l]{Percent\\Correct} & \makecell[l]{Percent\\Correct\\by Class} & Bias$^1$ & \makecell[l]{Critical\\Success\\Index$^2$} & \makecell[l]{Probability of\\False\\Detection$^3$} & \makecell[l]{Probability\\of\\Detection$^4$} & \makecell[l]{False\\Alarm\\Ratio$^5$} & \makecell[l]{Heidke\\Skill\\Score$^6$} & \makecell[l]{Peirce\\Skill\\Score$^7$} & \makecell[l]{Gerrity\\Score$^8$}\\
\midrule
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 1. Single-level/No opponent}}\\
\hspace{1em}No Injury & 50652 & 22504 & 150 &  & \textcolor{red}{69.07} & \textcolor{black}{\textbf{0.9293}} & \textcolor{black}{0.4988} & \textcolor{red}{0.2646} & \textcolor{black}{0.6421} & \textcolor{black}{0.309} &  &  & \\

\hspace{1em}Injury & 28232 & 62120 & 797 &  & \textcolor{red}{68.64} & \textcolor{black}{\textbf{1.0765}} & \textcolor{red}{0.5463} & \textcolor{black}{0.3636} & \textcolor{red}{0.7336} & \textcolor{red}{0.3185} &  &  & \\

Fatality & 2 & 51 & 3 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{68.55}} & \textcolor{black}{99.39} & \textcolor{black}{0.0589} & \textcolor{black}{0.003} & \textcolor{red}{3e-04} & \textcolor{black}{0.0032} & \textcolor{black}{0.9464} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3725}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3696}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.1902}}\\
\cmidrule{1-14}
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 2. Single-level/Opponent attributes}}\\
\hspace{1em}No Injury & 51531 & 17137 & 85 &  & \textcolor{black}{\textbf{72.9}} & \textcolor{black}{0.8715} & \textcolor{black}{\textbf{0.5362}} & \textcolor{black}{0.2011} & \textcolor{black}{\underline{0.6532}} & \textcolor{black}{\textbf{0.2505}} &  &  & \\

\hspace{1em}Injury & 27355 & 67514 & 864 &  & \textcolor{black}{\textbf{72.42}} & \textcolor{black}{1.1306} & \textcolor{black}{0.598} & \textcolor{black}{\underline{0.3535}} & \textcolor{black}{0.7973} & \textcolor{black}{0.2948} &  &  & \\

Fatality & 0 & 24 & 1 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{72.36}}} & \textcolor{black}{\underline{99.41}} & \textcolor{red}{0.0263} & \textcolor{red}{0.001} & \textcolor{black}{1e-04} & \textcolor{red}{0.0011} & \textcolor{red}{0.96} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4474}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4429}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.2265}}}\\
\cmidrule{1-14}
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 3. Hierarchical: Traffic unit}}\\
\hspace{1em}No Injury & 51103 & 17296 & 79 &  & \textcolor{black}{72.55} & \textcolor{black}{0.8681} & \textcolor{black}{0.5309} & \textcolor{black}{0.2029} & \textcolor{black}{0.6478} & \textcolor{black}{0.2537} &  &  & \\

\hspace{1em}Injury & 27783 & 67338 & 868 &  & \textcolor{black}{72.05} & \textcolor{black}{1.1336} & \textcolor{black}{0.5942} & \textcolor{black}{0.3589} & \textcolor{black}{0.7953} & \textcolor{black}{0.2985} &  &  & \\

Fatality & 0 & 41 & 3 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{72}} & \textcolor{black}{99.4} & \textcolor{black}{0.0463} & \textcolor{black}{0.003} & \textcolor{black}{3e-04} & \textcolor{black}{0.0032} & \textcolor{black}{0.9318} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.44}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.4357}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2239}}\\
\cmidrule{1-14}
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 4. Hierarchical: Opponent}}\\
\hspace{1em}No Injury & 51574 & 17316 & 84 &  & \textcolor{black}{\underline{72.82}} & \textcolor{black}{\underline{0.8744}} & \textcolor{black}{\underline{0.5356}} & \textcolor{black}{0.2032} & \textcolor{black}{\textbf{0.6538}} & \textcolor{black}{\underline{0.2523}} &  &  & \\

\hspace{1em}Injury & 27312 & 67336 & 863 &  & \textcolor{black}{\underline{72.33}} & \textcolor{black}{\underline{1.128}} & \textcolor{black}{0.5967} & \textcolor{black}{\textbf{0.3529}} & \textcolor{black}{0.7952} & \textcolor{black}{0.295} &  &  & \\

Fatality & 0 & 23 & 3 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{72.28}}} & \textcolor{black}{\textbf{99.41}} & \textcolor{black}{0.0274} & \textcolor{black}{0.0031} & \textcolor{black}{1e-04} & \textcolor{black}{0.0032} & \textcolor{black}{0.8846} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4458}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4414}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.2268}}}\\
\cmidrule{1-14}
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 1 Ensemble. Single-level/No opponent}}\\
\hspace{1em}No Injury & 34664 & 16434 & 63 &  & \textcolor{black}{69.88} & \textcolor{red}{0.8197} & \textcolor{red}{0.4393} & \textcolor{black}{0.1952} & \textcolor{red}{0.5554} & \textcolor{red}{0.3225} &  &  & \\

\hspace{1em}Injury & 27749 & 67120 & 829 &  & \textcolor{black}{69.35} & \textcolor{red}{1.1452} & \textcolor{black}{0.5985} & \textcolor{red}{0.4512} & \textcolor{black}{0.8032} & \textcolor{black}{0.2986} &  &  & \\

Fatality & 0 & 10 & 39 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.31}} & \textcolor{red}{99.39} & \textcolor{black}{0.0526} & \textcolor{black}{0.0414} & \textcolor{black}{\underline{1e-04}} & \textcolor{black}{0.0419} & \textcolor{black}{\underline{0.2041}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.3626}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.3521}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.201}}\\
\cmidrule{1-14}
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 2 Ensemble. Single-level/Opponent attributes}}\\
\hspace{1em}No Injury & 35443 & 16145 & 60 &  & \textcolor{black}{70.62} & \textcolor{black}{0.8275} & \textcolor{black}{0.4508} & \textcolor{black}{\textbf{0.1917}} & \textcolor{black}{0.5678} & \textcolor{black}{0.3138} &  &  & \\

\hspace{1em}Injury & 26974 & 67437 & 829 &  & \textcolor{black}{70.08} & \textcolor{black}{1.1393} & \textcolor{black}{\textbf{0.6054}} & \textcolor{black}{0.4389} & \textcolor{black}{\textbf{0.8067}} & \textcolor{black}{0.2919} &  &  & \\

Fatality & 0 & 13 & 42 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{70.05}} & \textcolor{black}{99.39} & \textcolor{black}{0.0591} & \textcolor{black}{0.0445} & \textcolor{black}{1e-04} & \textcolor{black}{0.0451} & \textcolor{black}{0.2364} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3784}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3678}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2106}}\\
\cmidrule{1-14}
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 3 Ensemble. Hierarchical: Traffic unit}}\\
\hspace{1em}No Injury & 35498 & 16204 & 60 &  & \textcolor{black}{70.62} & \textcolor{black}{0.8294} & \textcolor{black}{0.4512} & \textcolor{black}{\underline{0.1924}} & \textcolor{black}{0.5688} & \textcolor{black}{0.3142} &  &  & \\

\hspace{1em}Injury & 26913 & 67379 & 828 &  & \textcolor{black}{70.08} & \textcolor{black}{1.1379} & \textcolor{black}{\underline{0.6052}} & \textcolor{black}{0.4379} & \textcolor{black}{\underline{0.806}} & \textcolor{black}{\underline{0.2916}} &  &  & \\

Fatality & 0 & 13 & 45 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{70.04}} & \textcolor{black}{99.39} & \textcolor{black}{\textbf{0.0622}} & \textcolor{black}{\underline{0.0476}} & \textcolor{black}{1e-04} & \textcolor{black}{\underline{0.0482}} & \textcolor{black}{0.2241} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3786}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3681}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2123}}\\
\cmidrule{1-14}
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 4 Ensemble. Hierarchical: Opponent}}\\
\hspace{1em}No Injury & 35553 & 16297 & 59 &  & \textcolor{black}{70.59} & \textcolor{black}{0.8318} & \textcolor{black}{0.4514} & \textcolor{black}{0.1935} & \textcolor{black}{0.5697} & \textcolor{black}{0.3151} &  &  & \\

\hspace{1em}Injury & 26852 & 67271 & 827 &  & \textcolor{black}{70.06} & \textcolor{black}{1.1361} & \textcolor{black}{0.6046} & \textcolor{black}{0.437} & \textcolor{black}{0.8049} & \textcolor{black}{\textbf{0.2915}} &  &  & \\

Fatality & 0 & 10 & 46 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{70.02}} & \textcolor{black}{99.39} & \textcolor{black}{\underline{0.0601}} & \textcolor{black}{\textbf{0.0488}} & \textcolor{black}{\textbf{1e-04}} & \textcolor{black}{\textbf{0.0494}} & \textcolor{black}{\textbf{0.1786}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3783}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3679}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2127}}\\
\bottomrule
\multicolumn{14}{l}{\textit{Note: }}\\
\multicolumn{14}{l}{Bold numbers: best scores; underlined numbers: second best scores; red numbers: worst scores}\\
\multicolumn{14}{l}{\textsuperscript{1} $B>1$: class is overpredicted; $B<1$: class is underpredicted; }\\
\multicolumn{14}{l}{\textsuperscript{2} $CSI = 1$: perfect score; $CSI = 0$: no skill; }\\
\multicolumn{14}{l}{\textsuperscript{3} $F = 0$: perfect score; }\\
\multicolumn{14}{l}{\textsuperscript{4} $POD = 1$: perfect score; }\\
\multicolumn{14}{l}{\textsuperscript{5} $FAR = 0$: perfect score; }\\
\multicolumn{14}{l}{\textsuperscript{6} $HSS = 1$: perfect score; $HSS = 0$: no skill; $HSS < 0$: random is better; }\\
\multicolumn{14}{l}{\textsuperscript{7} $PSS = 1$: perfect score; $PSS = 0$: no skill; }\\
\multicolumn{14}{l}{\textsuperscript{8} $GS = 1$: perfect score; $GS = 0$: no skill.}\\
\end{tabular}}
\end{table}

\hypertarget{backcasting-verification-statistics}{%
\subsubsection{Backcasting: verification
statistics}\label{backcasting-verification-statistics}}

Table \ref{tab:backcast-outcomes-assessment} presents the results of the
verification exercise for the case of our out-of-sample predictions
(i.e., backasting). Qualitatively, the results are similar to those of
the nowcasting experiments, but with a somewhat weaker performance of
the ensemble models. This, again, supports the idea that these models
might be overfitting the process, as discussed in reference to the
aggregate forecasts (see Section \ref{sec:outcomes-shares}). Models 2
(full sample) and 4 (full sample) are again identified as the best
overall performers, and particularly Model 2 (full sample) performs
somewhat more adroitly with respect to Fatality in backcasting than it
did in nowcasting.

\begin{table}

\caption{\label{tab:backcast-outcomes-assessment}\label{tab:backcast-outcomes-assessment}Assessment of out-of-sample outcomes (backcasting using 2016 data set)}
\centering
\resizebox{\linewidth}{!}{
\fontsize{7}{9}\selectfont
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrllllllllll}
\toprule
\multicolumn{1}{c}{Predicted} & \multicolumn{3}{c}{Observed Outcome} & \multicolumn{10}{c}{Verification Statistics} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-14}
Outcome & No Injury & Injury & Fatality & \makecell[l]{Percent\\Correct} & \makecell[l]{Percent\\Correct\\by Class} & Bias$^1$ & \makecell[l]{Critical\\Success\\Index$^2$} & \makecell[l]{Probability of\\False\\Detection$^3$} & \makecell[l]{Probability\\of\\Detection$^4$} & \makecell[l]{False\\Alarm\\Ratio$^5$} & \makecell[l]{Heidke\\Skill\\Score$^6$} & \makecell[l]{Peirce\\Skill\\Score$^7$} & \makecell[l]{Gerrity\\Score$^8$}\\
\midrule
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 1. Single-level/No opponent}}\\
\hspace{1em}No Injury & 61683 & 27447 & 184 &  & \textcolor{red}{68.53} & \textcolor{black}{\textbf{0.9221}} & \textcolor{black}{0.4955} & \textcolor{red}{0.269} & \textcolor{black}{0.6368} & \textcolor{black}{0.3094} &  &  & \\

\hspace{1em}Injury & 35172 & 74073 & 915 &  & \textcolor{red}{68.12} & \textcolor{black}{\textbf{1.0842}} & \textcolor{red}{0.538} & \textcolor{black}{0.3684} & \textcolor{red}{0.729} & \textcolor{red}{0.3276} &  &  & \\

Fatality & 5 & 85 & 10 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{68.03}} & \textcolor{black}{99.4} & \textcolor{black}{0.0902} & \textcolor{black}{0.0083} & \textcolor{black}{5e-04} & \textcolor{black}{0.009} & \textcolor{red}{0.9} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3628}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3604}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1882}}\\
\cmidrule{1-14}
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 2. Single-level/Opponent attributes}}\\
\hspace{1em}No Injury & 62736 & 21013 & 106 &  & \textcolor{black}{\textbf{72.32}} & \textcolor{black}{0.8657} & \textcolor{black}{\textbf{0.5318}} & \textcolor{black}{0.2056} & \textcolor{black}{\underline{0.6477}} & \textcolor{black}{\textbf{0.2519}} &  &  & \\

\hspace{1em}Injury & 34124 & 80569 & 996 &  & \textcolor{black}{\textbf{71.86}} & \textcolor{black}{1.1386} & \textcolor{black}{0.5893} & \textcolor{black}{\underline{0.3585}} & \textcolor{black}{0.793} & \textcolor{black}{\textbf{0.3036}} &  &  & \\

Fatality & 0 & 23 & 7 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{71.81}}} & \textcolor{black}{\textbf{99.44}} & \textcolor{red}{0.0271} & \textcolor{black}{0.0062} & \textcolor{black}{\textbf{1e-04}} & \textcolor{black}{0.0063} & \textcolor{black}{\textbf{0.7667}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4373}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.4336}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\textbf{0.2241}}}\\
\cmidrule{1-14}
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 3. Hierarchical: Traffic unit}}\\
\hspace{1em}No Injury & 62248 & 21133 & 107 &  & \textcolor{black}{72.01} & \textcolor{black}{0.8619} & \textcolor{black}{0.5271} & \textcolor{black}{0.2068} & \textcolor{black}{0.6427} & \textcolor{black}{0.2544} &  &  & \\

\hspace{1em}Injury & 34610 & 80433 & 996 &  & \textcolor{black}{71.55} & \textcolor{black}{1.1421} & \textcolor{black}{0.5862} & \textcolor{black}{0.3634} & \textcolor{black}{0.7916} & \textcolor{black}{0.3068} &  &  & \\

Fatality & 2 & 39 & 6 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{71.5}} & \textcolor{black}{99.43} & \textcolor{black}{0.0424} & \textcolor{red}{0.0052} & \textcolor{black}{2e-04} & \textcolor{red}{0.0054} & \textcolor{black}{0.8723} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.431}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.4274}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.2205}}\\
\cmidrule{1-14}
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 4. Hierarchical: Opponent}}\\
\hspace{1em}No Injury & 62788 & 21247 & 102 &  & \textcolor{black}{\underline{72.23}} & \textcolor{black}{\underline{0.8686}} & \textcolor{black}{\underline{0.5312}} & \textcolor{black}{0.2078} & \textcolor{black}{\textbf{0.6482}} & \textcolor{black}{\underline{0.2537}} &  &  & \\

\hspace{1em}Injury & 34071 & 80331 & 1000 &  & \textcolor{black}{\underline{71.77}} & \textcolor{black}{\underline{1.1358}} & \textcolor{black}{0.5877} & \textcolor{black}{\textbf{0.358}} & \textcolor{black}{0.7906} & \textcolor{black}{0.3039} &  &  & \\

Fatality & 1 & 27 & 7 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{71.72}}} & \textcolor{black}{\underline{99.43}} & \textcolor{black}{0.0316} & \textcolor{black}{0.0062} & \textcolor{black}{\underline{1e-04}} & \textcolor{black}{0.0063} & \textcolor{black}{\underline{0.8}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4354}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.4318}}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{\underline{0.2233}}}\\
\cmidrule{1-14}
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 1 Ensemble. Single-level/No opponent}}\\
\hspace{1em}No Injury & 42896 & 20230 & 95 &  & \textcolor{black}{69.26} & \textcolor{red}{0.8162} & \textcolor{red}{0.4387} & \textcolor{black}{0.2011} & \textcolor{red}{0.5538} & \textcolor{red}{0.3215} &  &  & \\

\hspace{1em}Injury & 34546 & 79692 & 962 &  & \textcolor{black}{68.73} & \textcolor{red}{1.1519} & \textcolor{black}{0.588} & \textcolor{red}{0.4522} & \textcolor{black}{0.7968} & \textcolor{black}{0.3082} &  &  & \\

Fatality & 15 & 91 & 15 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{68.67}} & \textcolor{black}{99.35} & \textcolor{black}{0.1129} & \textcolor{black}{0.0127} & \textcolor{black}{6e-04} & \textcolor{black}{0.014} & \textcolor{black}{0.876} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.3539}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.3447}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{red}{0.1831}}\\
\cmidrule{1-14}
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 2 Ensemble. Single-level/Opponent attributes}}\\
\hspace{1em}No Injury & 43485 & 19937 & 95 &  & \textcolor{black}{69.76} & \textcolor{black}{0.82} & \textcolor{black}{0.446} & \textcolor{black}{\textbf{0.1981}} & \textcolor{black}{0.5614} & \textcolor{black}{0.3154} &  &  & \\

\hspace{1em}Injury & 33954 & 80009 & 961 &  & \textcolor{black}{69.23} & \textcolor{black}{1.1487} & \textcolor{black}{\textbf{0.5928}} & \textcolor{black}{0.4446} & \textcolor{black}{\textbf{0.7997}} & \textcolor{black}{\underline{0.3038}} &  &  & \\

Fatality & 20 & 103 & 15 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.16}} & \textcolor{black}{99.34} & \textcolor{black}{\underline{0.1289}} & \textcolor{black}{0.0126} & \textcolor{black}{7e-04} & \textcolor{black}{0.014} & \textcolor{black}{0.8913} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3644}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.355}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1883}}\\
\cmidrule{1-14}
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 3 Ensemble. Hierarchical: Traffic unit}}\\
\hspace{1em}No Injury & 43526 & 20033 & 92 &  & \textcolor{black}{69.73} & \textcolor{black}{0.8217} & \textcolor{black}{0.446} & \textcolor{black}{\underline{0.199}} & \textcolor{black}{0.5619} & \textcolor{black}{0.3162} &  &  & \\

\hspace{1em}Injury & 33915 & 79915 & 964 &  & \textcolor{black}{69.19} & \textcolor{black}{1.1474} & \textcolor{black}{\underline{0.5923}} & \textcolor{black}{0.4441} & \textcolor{black}{\underline{0.7988}} & \textcolor{black}{0.3038} &  &  & \\

Fatality & 18 & 102 & 16 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.13}} & \textcolor{black}{99.34} & \textcolor{black}{0.1269} & \textcolor{black}{\underline{0.0134}} & \textcolor{black}{7e-04} & \textcolor{black}{\underline{0.0149}} & \textcolor{black}{0.8824} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3639}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3546}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1885}}\\
\cmidrule{1-14}
\addlinespace[0.3em]
\multicolumn{14}{l}{\textbf{Model 4 Ensemble. Hierarchical: Opponent}}\\
\hspace{1em}No Injury & 43561 & 20160 & 94 &  & \textcolor{black}{69.67} & \textcolor{black}{0.8238} & \textcolor{black}{0.4458} & \textcolor{black}{0.2003} & \textcolor{black}{0.5624} & \textcolor{black}{0.3174} &  &  & \\

\hspace{1em}Injury & 33875 & 79762 & 959 &  & \textcolor{black}{69.14} & \textcolor{black}{1.1456} & \textcolor{black}{0.5914} & \textcolor{black}{0.4436} & \textcolor{black}{0.7974} & \textcolor{black}{0.304} &  &  & \\

Fatality & 25 & 107 & 17 & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{69.07}} & \textcolor{red}{99.34} & \textcolor{black}{\textbf{0.1393}} & \textcolor{black}{\textbf{0.0141}} & \textcolor{red}{7e-04} & \textcolor{black}{\textbf{0.0159}} & \textcolor{black}{0.8859} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3629}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.3538}} & \multirow{-3}{*}{\raggedright\arraybackslash \textcolor{black}{0.1886}}\\
\bottomrule
\multicolumn{14}{l}{\textit{Note: }}\\
\multicolumn{14}{l}{Bold numbers: best scores; underlined numbers: second best scores; red numbers: worst scores}\\
\multicolumn{14}{l}{\textsuperscript{1} $B>1$: class is overpredicted; $B<1$: class is underpredicted; }\\
\multicolumn{14}{l}{\textsuperscript{2} $CSI = 1$: perfect score; $CSI = 0$: no skill; }\\
\multicolumn{14}{l}{\textsuperscript{3} $F = 0$: perfect score; }\\
\multicolumn{14}{l}{\textsuperscript{4} $POD = 1$: perfect score; }\\
\multicolumn{14}{l}{\textsuperscript{5} $FAR = 0$: perfect score; }\\
\multicolumn{14}{l}{\textsuperscript{6} $HSS = 1$: perfect score; $HSS = 0$: no skill; $HSS < 0$: random is better; }\\
\multicolumn{14}{l}{\textsuperscript{7} $PSS = 1$: perfect score; $PSS = 0$: no skill; }\\
\multicolumn{14}{l}{\textsuperscript{8} $GS = 1$: perfect score; $GS = 0$: no skill.}\\
\end{tabular}}}
\end{table}

\hypertarget{sec:further-considerations}{%
\section{Further considerations}\label{sec:further-considerations}}

As discussed in Section \ref{sec:methods}, there is a rich selection of
modelling approaches that are applicable to crash severity analysis.
Based on the literature, we limited our empirical assessment of
modelling strategies to only one model, namely the ordinal logit. On the
other hand, since the modelling strategies discussed here all relate to
the specification of the latent function and data subsetting, it is a
relatively simple matter to extend them to other modelling approaches.
For example, take Expression
\ref{eq:latent-function-with-opponent-variables} and add a random
component \(\mu_{k}\) as follows:

\begin{equation}
\label{eq:latent-function-with-opponent-variables-and-random-component}
y_{itk}^*=\sum_{l=1}^L\alpha_lp_{itkl} + \sum_{m=1}^M\beta_mu_{tkm} + \sum_{q=1}^Q\kappa_qc_{kq} + \sum_{r=1}^R\delta_ro_{jvkr} + \mu_{k} + \epsilon_{itk}
\end{equation}

The addition of the random component in this fashion would help to
capture, when appropriate, unobserved heterogeneity at the level of the
crash (this is similar to the random intercepts approach in multi-level
modelling; also see Mannering et al., 2016). As a second example, take
Expressions \ref{eq:latent-function-with-hierarchical-traffic-unit} to
\ref{eq:hierarchical-traffic-unit-coefficients} and add a random
component to a hierarchical coefficient, to obtain:

\begin{equation}
\label{eq:hierarchical-traffic-unit-coefficients-with-random-component}
\begin{array}{rcl}\
\beta_{m}u_{tkm} &=& \big( \beta_{m1} + \beta_{m2}p_{itk2} + \cdots + \beta_{mL}p_{itkL} + \mu_{mk}\big)u_{tkm}\\ 
&=& \beta_{m1}u_{tkm} + \beta_{m2}p_{itk2}u_{tkm} + \cdots + \beta_{mL}p_{itkL}u_{tkm} + \mu_{mk}u_{tkm}
\end{array}
\end{equation}

This is similar to the random slopes strategy in multi-level modelling.

We do not report results regarding other modelling strategies. On the
one hand, more sophisticated modelling frameworks are generally capable
of improving the performance of a model. On the other hand, there are
well-known challenges in the estimation of more sophisticated models
(see Lenguerrand et al., 2006, p. 47, for a discussion of convergence
issues in models with mixed effects; Mannering et al., 2016, p. 13, for
some considerations regarding the complexity and cost of estimating more
complex models; and Bogue et al., 2017, p. 27, on computational demands
of models with random components). The additional cost and complexity of
more sophisticated modelling approaches would, in our view, have greatly
complicated our empirical assessment, particularly considering the large
size of the sample involved in this research (a data set with over
164,000 records in the case of the full sample models). That said, we
experimented with a model with random components using monthly subsets
of data to find that, indeed, estimation takes considerably longer, is
more demanding in terms of fixing potential estimation quirks, and in
the end resulted in variance components that could not be reliably
estimated as different from zero (results can be consulted in the source
R Notebook). For this reason, we choose to leave the application of more
sophisticated models as a matter for future research.

\hypertarget{sec:concluding-remarks}{%
\section{Concluding remarks}\label{sec:concluding-remarks}}

The study of crash severity is an important component of accident
research, as seen from a large and vibrant literature and numerous
applications. Part of this literature covers different modelling
strategies that can be used to model complex hierarchical, multievent
outcomes such as the severity of injuries following a collision. In this
paper, our objective has been to assess the performance of different
strategies to model opponent effects in two-vehicle crashes. In broad
terms, three strategies were considered: 1) incorporating opponent-level
variables in the model; 2) single- versus multi-level model
specifications; and 3) sample subsetting and estimation of separate
models for different types of individual-opponent interactions. The
empirical evaluation was based on data from Canada's National Crash
Database and the application of ordered probit models. A suite of models
that implemented the various strategies considered was estimated using
data from 2017. We then assessed the performance of the models using one
information criterion (AIC). Furthermore, the predictive performance of
the models was assessed in terms of both nowcasting (in-sample
predictions) and backcasting (out-of-sample predictions), the latter
using data from 2016.

The results of the empirical assessment strongly suggest that
incorporating opponent effects can greatly improve the goodness-of-fit
and predictive performance of a model. Two modelling strategies appear
to outperform the rest: a relatively simple single-level modelling
approach that incorporates opponent effects, and a hierarchical
modelling approach with nested opponent effects. There was some evidence
that subsetting the sample can improve the results in some cases (e.g.,
when modelling the severity of crashes involving active travelers or
motorcyclists), but possibly at the risk of overfitting the process and
leading to worse out-of-sample performance. In this paper we did not
compare individual models in our ensemble approach, but we suggest that
this is an avenue for future research.

The results of this research should be informative to analysts
interested in crashes involving two parties, since it provides some
useful guidelines regarding the specification of opponent effects. Not
only do opponent effects improve the goodness of fit and performance of
models, they also add rich insights into their effects. The focus on
this paper was on performance, and for space reasons it is not possible
to include an examination of the best model without failing to do it
justice. We plan to report the results of the best-fitting model in a
future paper.

The analysis also opens up a few avenues for future research. First, for
reasons discussed in Section \ref{sec:further-considerations}, we did
not consider more sophisticated modelling approaches, such as models
with random components, partial proportional odds, ranked ordered
models, or multinomial models, to mention just a few possibilities.
Secondly, we only considered the performance of the models when making
predictions for the full sample. That is, the submodels in the ensembles
were not compared in detail, just their aggregate results when
predicting the full sample. However, the goodness-of-fit was not
uniformly better for any one modelling strategy when the data were
subset, and it is possible that individual models perform better for a
certain subset than competitors that are part of a better ensemble,
overall. For this reason, we suggest that additional work with ensemble
approaches is warranted. Finally, it is clear that the models do not
generally do well when predicting the least frequent class of outcome,
namely Fatality. It would be worthwhile to further investigate
approaches for so-called imbalanced learning, a task where Torrao et
al.~(2014) have already made some headway.

Finally, as an aside, this paper is, to the best of our knowledge, the
first example of reproducible research in crash severity analysis. By
providing the data \emph{and} code for the analysis, it is our hope that
this will allow other researchers to easily verify the results, and to
extend them. A common practice in the machine learning community is to
use canonical data set to demonstrate the performance of new techniques.
Sharing code and data has remained relatively rare in transportation
research, and we would like to suggest that the data sets used in this
research could constitute one such canonical data set for future
methodological developments.

\hypertarget{acknowledgments}{%
\section*{Acknowledgments}\label{acknowledgments}}
\addcontentsline{toc}{section}{Acknowledgments}

This research was supported by a Research Excellence grant from McMaster
University. The authors wish to express their gratitude to the Office of
the Provost, the Faculty of Science, and the Faculty of Engineering for
their generous support. The comments of four anonymous reviewers are
also duly acknowledged: their feedback helped to improve the quality and
clarity of presentation of this paper.

\hypertarget{appendix}{%
\section*{Appendix}\label{appendix}}
\addcontentsline{toc}{section}{Appendix}

Variable definitions in Canada's National Collision Database.

\begingroup\fontsize{7}{9}\selectfont

\begin{longtable}[t]{ll>{\raggedright\arraybackslash}p{32em}}
\caption{\label{tab:ncdb-descriptives-collision}\label{tab:ncdb-descriptives-collision}Contents of National Collision Database: Collision-level variables}\\
\toprule
Variable & Description & Notes\\
\midrule
\endfirsthead
\caption[]{\label{tab:ncdb-descriptives-collision}Contents of National Collision Database: Collision-level variables \textit{(continued)}}\\
\toprule
Variable & Description & Notes\\
\midrule
\endhead
\
\endfoot
\bottomrule
\multicolumn{3}{l}{\textit{Note: }}\\
\multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/data set/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
\multicolumn{3}{l}{Source data files for analysis also available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
\endlastfoot
\rowcolor{gray!6}  C\_CASE & Unique collision identifier & Unique identifier for collisions\\
C\_YEAR & Year & Last two digits of year.\\
\rowcolor{gray!6}  C\_MNTH & Month & 14 levels: January - December; unknown; not reported by jurisdiction.\\
C\_WDAY & Day of week & 9 levels: Monday - Sunday; unknown; not reported by jurisdiction.\\
\rowcolor{gray!6}  C\_HOUR & Collision hour & 25 levels: hourly intervals; unknown; not reported by jurisdiction.\\
\addlinespace
C\_SEV & Collision severity & 4 levels: collision producing at least one fatality; collision producing non-fatal injury; unknown; not reported by jurisdiction.\\
\rowcolor{gray!6}  C\_VEHS & Number of vehicles involved in collision & Number of vehicles: 1-98 vehicles involved; 99 or more vehicles involved; unknown; not reported by jurisdiction.\\
C\_CONF & Collision configuration & 21 levels: SINGLE VEHICLE: Hit a moving object (e.g. a person or an animal); Hit a stationary object (e.g. a tree); Ran off left shoulder; Ran off right shoulder; Rollover on roadway; Any other single vehicle collision configuration; TWO-VEHICLES SAME DIRECTION OF TRAVEL: Rear-end collision; Side swipe; One vehicle passing to the left of the other, or left turn conflict; One vehicle passing to the right of the other, or right turn conflict; Any other two vehicle - same direction of travel configuration; TWO-VEHICLES DIFFERENT DIRECTION OF TRAVEL: Head-on collision; Approaching side-swipe; Left turn across opposing traffic; Right turn, including turning conflicts; Right angle collision; Any other two-vehicle - different direction of travel configuration; TWO-VEHICLES, HIT A PARKED VEHICLE: Hit a parked motor vehicle; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
\rowcolor{gray!6}  C\_RCFG & Roadway configuration & 15 levels: Non-intersection; At an intersection of at least two public roadways; Intersection with parking lot entrance/exit, private driveway or laneway; Railroad level crossing; Bridge, overpass, viaduct; Tunnel or underpass; Passing or climbing lane; Ramp; Traffic circle; Express lane of a freeway system; Collector lane of a freeway system; Transfer lane of a freeway system; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
C\_WTHR & Weather condition & 10 levels: Clear and sunny; Overcast, cloudy but no precipitation; Raining; Snowing, not including drifting snow; Freezing rain, sleet, hail; Visibility limitation; Strong wind; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
\addlinespace
\rowcolor{gray!6}  C\_RSUR & Road surface & 12 levels: Dry, normal; Wet; Snow (fresh, loose snow); Slush, wet snow; Icy, packed snow; Debris on road (e.g., sand/gravel/dirt); Muddy; Oil; Flooded; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
C\_RALN & Road alignment & 9 levels: Straight and level; Straight with gradient; Curved and level; Curved with gradient; Top of hill or gradient; Bottom of hill or gradient; Choice is other than the preceding values; unknown;not reported by jurisdiction.\\
\rowcolor{gray!6}  C\_TRAF & Traffic control & 21 levels: Traffic signals fully operational; Traffic signals in flashing mode; Stop sign; Yield sign; Warning sign; Pedestrian crosswalk; Police officer; School guard, flagman; School crossing; Reduced speed zone; No passing zone sign; Markings on the road; School bus stopped with school bus signal lights flashing; School bus stopped with school bus signal lights not flashing; Railway crossing with signals, or signals and gates; Railway crossing with signs only; Control device not specified; No control present; Choice is other than the preceding values; unknown; not reported by jurisdiction.\\*
\end{longtable}
\endgroup{}

\begingroup\fontsize{7}{9}\selectfont

\begin{longtable}[t]{ll>{\raggedright\arraybackslash}p{32em}}
\caption{\label{tab:ncdb-descriptives-vehicle}\label{tab:ncdb-descriptives-vehicle}Contents of National Collision Database: Traffic unit-level variables}\\
\toprule
Variable & Description & Notes\\
\midrule
\endfirsthead
\caption[]{\label{tab:ncdb-descriptives-vehicle}Contents of National Collision Database: Traffic unit-level variables \textit{(continued)}}\\
\toprule
Variable & Description & Notes\\
\midrule
\endhead
\
\endfoot
\bottomrule
\multicolumn{3}{l}{\textit{Note: }}\\
\multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/data set/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
\multicolumn{3}{l}{Source data files for analysis also available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
\endlastfoot
\rowcolor{gray!6}  V\_ID & Vehicle sequence number & Number of vehicles: 1-98; Pedestrian sequence number: 99; unknown.\\
V\_TYPE & Vehicle type & 21 levels: Light Duty Vehicle (Passenger car, Passenger van, Light utility vehicles and light duty pick up trucks); Panel/cargo van (<= 4536 KG GVWR   Panel or window type of van designed primarily for carrying goods); Other trucks and vans (<= 4536 KG GVWR); Unit trucks (> 4536 KG GVWR); Road tractor; School bus; Smaller school bus (< 25 passengers); Urban and Intercity Bus; Motorcycle and moped; Off road vehicles; Bicycle; Purpose-built motorhome; Farm equipment; Construction equipment; Fire engine; Snowmobile; Street car; Data element is not applicable  (e.g. dummy vehicle record created for pedestrian); Choice is other than the preceding values; unknown; not reported by jurisdiction.\\
\rowcolor{gray!6}  V\_YEAR & Vehicle model year & Model year; dummy for pedestrians; unknown; not reported by jurisdiction.\\*
\end{longtable}
\endgroup{}

\begingroup\fontsize{7}{9}\selectfont

\begin{longtable}[t]{ll>{\raggedright\arraybackslash}p{32em}}
\caption{\label{tab:ncdb-descriptives-person}\label{tab:ncdb-descriptives-person}Contents of National Collision Database: Personal-level variables}\\
\toprule
Variable & Description & Notes\\
\midrule
\endfirsthead
\caption[]{\label{tab:ncdb-descriptives-person}Contents of National Collision Database: Personal-level variables \textit{(continued)}}\\
\toprule
Variable & Description & Notes\\
\midrule
\endhead
\
\endfoot
\bottomrule
\multicolumn{3}{l}{\textit{Note: }}\\
\multicolumn{3}{l}{Source NCDB available from https://open.canada.ca/data/en/data set/1eb9eba7-71d1-4b30-9fb1-30cbdab7e63a}\\
\multicolumn{3}{l}{Source data files for analysis also available from https://drive.google.com/open?id=12aJtVBaQ4Zj0xa7mtfqxh0E48hKCb\_XV}\\
\endlastfoot
\rowcolor{gray!6}  P\_ID & Person sequence number & Sequence number: 1-99; Not applicable (dummy for parked vehicles); not reported by jurisdiction.\\
P\_SEX & Person sex & 5 levels: Male; Female; Not applicable (dummy for parked vehicles); unknown (runaway vehicle); not reported by jurisdiction.\\
\rowcolor{gray!6}  P\_AGE & Person age & Age: less than 1 year; 1-98 years old; 99 years or older; Not applicable (dummy for parked vehicles); unknown (runaway vehicle); not reported by jurisdiction.\\
P\_PSN & Person position & Person position: Driver; Passenger front row, center; Passenger front row, right outboard (including motorcycle passenger in sidecar); Passenger second row, left outboard, including motorcycle passenger; Passenger second row, center; Passenger second row, right outboard; Passenger third row, left outboard;...; Position unknown, but the person was definitely an occupant; Sitting on someone’s lap; Outside passenger compartment; Pedestrian; Not applicable (dummy for parked vehicles); Choice is other than the preceding values; unknown (runaway vehicle); not reported by jurisdiction.\\
\rowcolor{gray!6}  P\_ISEV & Medical treatment required & 6 levels: No Injury; Injury; Fatality; Not applicable (dummy for parked vehicles); Choice is other than the preceding values; unknown (runaway vehicle); not reported by jurisdiction.\\
\addlinespace
P\_SAFE & Safety device used & 11 levels: No safety device used; Safety device used; Helmet worn; Reflective clothing worn; Both helmet and reflective clothing used; Other safety device used; No safety device equipped   (e.g. buses); Not applicable (dummy for parked vehicles); Choice is other than the preceding values; unknown (runaway vehicle); not reported by jurisdiction.\\
\rowcolor{gray!6}  P\_USER & Road user class & 6 levels: Motor Vehicle Driver; Motor Vehicle Passenger; Pedestrian; Bicyclist; Motorcyclist; Not stated/Other/Unknown.\\*
\end{longtable}
\endgroup{}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Amoh2017effect}{}%
Amoh-Gyimah, R., Aidoo, E.N., Akaateba, M.A., Appiah, S.K., 2017. The
effect of natural and built environmental characteristics on
pedestrian-vehicle crash severity in ghana. International Journal of
Injury Control and Safety Promotion 24, 459--468.
doi:\href{https://doi.org/10.1080/17457300.2016.1232274}{10.1080/17457300.2016.1232274}

\leavevmode\hypertarget{ref-Aziz2013exploring}{}%
Aziz, H.M.A., Ukkusuri, S.V., Hasan, S., 2013. Exploring the
determinants of pedestrian-vehicle crash severity in new york city.
Accident Analysis and Prevention 50, 1298--1309.
doi:\href{https://doi.org/10.1016/j.aap.2012.09.034}{10.1016/j.aap.2012.09.034}

\leavevmode\hypertarget{ref-Begueria2006validation}{}%
Beguería, S., 2006. Validation and evaluation of predictive models in
hazard assessment and risk management. Natural Hazards 37, 315--329.
doi:\href{https://doi.org/10.1007/s11069-005-5182-6}{10.1007/s11069-005-5182-6}

\leavevmode\hypertarget{ref-Bogue2017modified}{}%
Bogue, S., Paleti, R., Balan, L., 2017. A modified rank ordered logit
model to analyze injury severity of occupants in multivehicle crashes.
Analytic Methods in Accident Research 14, 22--40.
doi:\href{https://doi.org/10.1016/j.amar.2017.03.001}{10.1016/j.amar.2017.03.001}

\leavevmode\hypertarget{ref-Casetti1972generating}{}%
Casetti, E., 1972. Generating models by the expansion method:
Applications to geographic research. Geographical Analysis 4, 81--91.

\leavevmode\hypertarget{ref-Chang2006analysis}{}%
Chang, L.Y., Wang, H.W., 2006. Analysis of traffic injury severity: An
application of non-parametric classification tree techniques. Accident
Analysis and Prevention 38, 1019--1027.
doi:\href{https://doi.org/10.1016/j.aap.2006.04.009}{10.1016/j.aap.2006.04.009}

\leavevmode\hypertarget{ref-Chen2019investigation}{}%
Chen, F., Song, M.T., Ma, X.X., 2019. Investigation on the injury
severity of drivers in rear-end collisions between cars using a random
parameters bivariate ordered probit model. International Journal of
Environmental Research and Public Health 16.
doi:\href{https://doi.org/10.3390/ijerph16142632}{10.3390/ijerph16142632}

\leavevmode\hypertarget{ref-Chiou2013modeling}{}%
Chiou, Y.C., Hwang, C.C., Chang, C.C., Fu, C., 2013. Modeling
two-vehicle crash severity by a bivariate generalized ordered probit
approach. Accident Analysis and Prevention 51, 175--184.
doi:\href{https://doi.org/10.1016/j.aap.2012.11.008}{10.1016/j.aap.2012.11.008}

\leavevmode\hypertarget{ref-Dissanayake2002factors}{}%
Dissanayake, S., Lu, J.J., 2002. Factors influential in making an injury
severity difference to older drivers involved in fixed object-passenger
car crashes. Accident Analysis and Prevention 34, 609--618.
doi:\href{https://doi.org/10.1016/s0001-4575(01)00060-4}{10.1016/s0001-4575(01)00060-4}

\leavevmode\hypertarget{ref-Duddu2018modeling}{}%
Duddu, V.R., Penmetsa, P., Pulugurtha, S.S., 2018. Modeling and
comparing injury severity of at-fault and not at-fault drivers in
crashes. Accident Analysis and Prevention 120, 55--63.
doi:\href{https://doi.org/10.1016/j.aap.2018.07.036}{10.1016/j.aap.2018.07.036}

\leavevmode\hypertarget{ref-Effati2015geospatial}{}%
Effati, M., Thill, J.C., Shabani, S., 2015. Geospatial and machine
learning techniques for wicked social science problems: Analysis of
crash severity on a regional highway corridor. Journal of Geographical
Systems 17, 107--135.
doi:\href{https://doi.org/10.1007/s10109-015-0210-x}{10.1007/s10109-015-0210-x}

\leavevmode\hypertarget{ref-Gong2017modeling}{}%
Gong, L.F., Fan, W.D., 2017. Modeling single-vehicle run-off-road crash
severity in rural areas: Accounting for unobserved heterogeneity and age
difference. Accident Analysis and Prevention 101, 124--134.
doi:\href{https://doi.org/10.1016/j.aap.2017.02.014}{10.1016/j.aap.2017.02.014}

\leavevmode\hypertarget{ref-Haleem2013effect}{}%
Haleem, K., Gan, A., 2013. Effect of driver's age and side of impact on
crash severity along urban freeways: A mixed logit approach. Journal of
Safety Research 46, 67--76.
doi:\href{https://doi.org/10.1016/j.jsr.2013.04.002}{10.1016/j.jsr.2013.04.002}

\leavevmode\hypertarget{ref-Hedeker1994random}{}%
Hedeker, D., Gibbons, R.D., 1994. A random-effects ordinal
regression-model for multilevel analysis. Biometrics 50, 933--944.

\leavevmode\hypertarget{ref-Iranitalab2017comparison}{}%
Iranitalab, A., Khattak, A., 2017. Comparison of four statistical and
machine learning methods for crash severity prediction. Accident
Analysis and Prevention 108, 27--36.
doi:\href{https://doi.org/10.1016/j.aap.2017.08.008}{10.1016/j.aap.2017.08.008}

\leavevmode\hypertarget{ref-Islam2014comprehensive}{}%
Islam, S., Jones, S.L., Dye, D., 2014. Comprehensive analysis of single-
and multi-vehicle large truck at-fault crashes on rural and urban
roadways in alabama. Accident Analysis \& Prevention 67, 148--158.
doi:\href{https://doi.org/https://doi.org/10.1016/j.aap.2014.02.014}{https://doi.org/10.1016/j.aap.2014.02.014}

\leavevmode\hypertarget{ref-Khan2015exploring}{}%
Khan, G., Bill, A.R., Noyce, D.A., 2015. Exploring the feasibility of
classification trees versus ordinal discrete choice models for analyzing
crash severity. Transportation Research Part C-Emerging Technologies 50,
86--96.
doi:\href{https://doi.org/10.1016/j.trc.2014.10.003}{10.1016/j.trc.2014.10.003}

\leavevmode\hypertarget{ref-Kim2013driver}{}%
Kim, J.K., Ulfarsson, G.F., Kim, S., Shankar, V.N., 2013. Driver-injury
severity in single-vehicle crashes in california: A mixed logit analysis
of heterogeneity due to age and gender. Accident Analysis and Prevention
50, 1073--1081.
doi:\href{https://doi.org/10.1016/j.aap.2012.08.011}{10.1016/j.aap.2012.08.011}

\leavevmode\hypertarget{ref-Kim1995personal}{}%
Kim, K., Nitz, L., Richardson, J., Li, L., 1995. PERSONAL and behavioral
predictors of automobile crash and injury severity. Accident Analysis
and Prevention 27, 469--481.
doi:\href{https://doi.org/10.1016/0001-4575(95)00001-g}{10.1016/0001-4575(95)00001-g}

\leavevmode\hypertarget{ref-Lee2014analysis}{}%
Lee, C., Li, X.C., 2014. Analysis of injury severity of drivers involved
in single- and two-vehicle crashes on highways in ontario. Accident
Analysis and Prevention 71, 286--295.
doi:\href{https://doi.org/10.1016/j.aap.2014.06.008}{10.1016/j.aap.2014.06.008}

\leavevmode\hypertarget{ref-Lenguerrand2006modelling}{}%
Lenguerrand, E., Martin, J.L., Laumon, B., 2006. Modelling the
hierarchical structure of road crash data - application to severity
analysis. Accident Analysis and Prevention 38, 43--53.
doi:\href{https://doi.org/10.1016/j.aap.2005.06.021}{10.1016/j.aap.2005.06.021}

\leavevmode\hypertarget{ref-Li2017interplay}{}%
Li, L., Hasnine, M.S., Habib, K.M.N., Persaud, B., Shalaby, A., 2017.
Investigating the interplay between the attributes of at-fault and
not-at-fault drivers and the associated impacts on crash injury
occurrence and severity level. Journal of Transportation Safety \&
Security 9, 439--456.
doi:\href{https://doi.org/10.1080/19439962.2016.1237602}{10.1080/19439962.2016.1237602}

\leavevmode\hypertarget{ref-Ma2008multivariate}{}%
Ma, J.M., Kockelman, K.M., Damien, P., 2008. A multivariate
poisson-lognormal regression model for prediction of crash counts by
severity, using bayesian methods. Accident Analysis and Prevention 40,
964--975.
doi:\href{https://doi.org/10.1016/j.aap.2007.11.002}{10.1016/j.aap.2007.11.002}

\leavevmode\hypertarget{ref-Maddala1986limited}{}%
Maddala, G.S., 1986. Limited-dependent and qualitative variables in
econometrics. Cambridge university press.

\leavevmode\hypertarget{ref-Mannering2016unobserved}{}%
Mannering, F., Shankar, V., Bhat, C.R., 2016. Unobserved heterogeneity
and the statistical analysis of highway accident data. Analytic Methods
in Accident Research 11, 1--16.
doi:\href{https://doi.org/10.1016/j.amar.2016.04.001}{10.1016/j.amar.2016.04.001}

\leavevmode\hypertarget{ref-Montella2013crash}{}%
Montella, A., Andreassen, D., Tarko, A.P., Turner, S., Mauriello, F.,
Imbriani, L.L., Romero, M.A., 2013. Crash databases in australasia, the
european union, and the united states review and prospects for
improvement. Transportation Research Record 128--136.
doi:\href{https://doi.org/10.3141/2386-15}{10.3141/2386-15}

\leavevmode\hypertarget{ref-Mooradian2013analysis}{}%
Mooradian, J., Ivan, J.N., Ravishanker, N., Hu, S., 2013. Analysis of
driver and passenger crash injury severity using partial proportional
odds models. Accident Analysis and Prevention 58, 53--58.
doi:\href{https://doi.org/10.1016/j.aap.2013.04.022}{10.1016/j.aap.2013.04.022}

\leavevmode\hypertarget{ref-Osman2018injury}{}%
Osman, M., Mishra, S., Paleti, R., 2018. Injury severity analysis of
commercially-licensed drivers in single-vehicle crashes: Accounting for
unobserved heterogeneity and age group differences. Accident Analysis
and Prevention 118, 289--300.
doi:\href{https://doi.org/10.1016/j.aap.2018.05.004}{10.1016/j.aap.2018.05.004}

\leavevmode\hypertarget{ref-Penmetsa2017examining}{}%
Penmetsa, P., Pulugurtha, S.S., Duddu, V.R., 2017. Examining injury
severity of not-at-fault drivers in two-vehicle crashes. Transportation
Research Record 164--173.
doi:\href{https://doi.org/10.3141/2659-18}{10.3141/2659-18}

\leavevmode\hypertarget{ref-Provost1998glossary}{}%
Provost, F., Kohavi, R., 1998. Glossary of terms. Journal of Machine
Learning 30, 271--274.

\leavevmode\hypertarget{ref-Rana2010copula}{}%
Rana, T.A., Sikder, S., Pinjari, A.R., 2010. Copula-based method for
addressing endogeneity in models of severity of traffic crash injuries
application to two-vehicle crashes. Transportation Research Record
75--87. doi:\href{https://doi.org/10.3141/2147-10}{10.3141/2147-10}

\leavevmode\hypertarget{ref-Rifaat2007accident}{}%
Rifaat, S.M., Chin, H.C., 2007. Accident severity analysis using ordered
probit model. Journal of Advanced Transportation 41, 91--114.
doi:\href{https://doi.org/10.1002/atr.5670410107}{10.1002/atr.5670410107}

\leavevmode\hypertarget{ref-Roorda2010trip}{}%
Roorda, M.J., Paez, A., Morency, C., Mercado, R., Farber, S., 2010. Trip
generation of vulnerable populations in three canadian cities: A spatial
ordered probit approach. Transportation 37, 525--548.
doi:\href{https://doi.org/10.1007/s11116-010-9263-3}{10.1007/s11116-010-9263-3}

\leavevmode\hypertarget{ref-Sasidharan2014partial}{}%
Sasidharan, L., Menendez, M., 2014. Partial proportional odds model-an
alternate choice for analyzing pedestrian crash injury severities.
Accident Analysis and Prevention 72, 330--340.
doi:\href{https://doi.org/10.1016/j.aap.2014.07.025}{10.1016/j.aap.2014.07.025}

\leavevmode\hypertarget{ref-Savolainen2007probabilistic}{}%
Savolainen, P., Mannering, F., 2007. Probabilistic models of
motorcyclists' injury severities in single- and multi-vehicle crashes.
Accident Analysis and Prevention 39, 955--963.
doi:\href{https://doi.org/10.1016/j.aap.2006.12.016}{10.1016/j.aap.2006.12.016}

\leavevmode\hypertarget{ref-Savolainen2011statistical}{}%
Savolainen, P.T., Mannering, F., Lord, D., Quddus, M.A., 2011. The
statistical analysis of highway crash-injury severities: A review and
assessment of methodological alternatives. Accident Analysis and
Prevention 43, 1666--1676.
doi:\href{https://doi.org/10.1016/j.aap.2011.03.025}{10.1016/j.aap.2011.03.025}

\leavevmode\hypertarget{ref-Shamsunnahar2014examining}{}%
Shamsunnahar, Y., Eluru, N., Pinjari, A.R., Tay, R., 2014. Examining
driver injury severity in two vehicle crashes - a copula based approach.
Accident Analysis and Prevention 66, 120--135.
doi:\href{https://doi.org/10.1016/j.aap.2014.01.018}{10.1016/j.aap.2014.01.018}

\leavevmode\hypertarget{ref-Tang2019crash}{}%
Tang, J.J., Liang, J., Han, C.Y., Li, Z.B., Huang, H.L., 2019. Crash
injury severity analysis using a two-layer stacking framework. Accident
Analysis and Prevention 122, 226--238.
doi:\href{https://doi.org/10.1016/j.aap.2018.10.016}{10.1016/j.aap.2018.10.016}

\leavevmode\hypertarget{ref-Tay2011multinomial}{}%
Tay, R., Choi, J., Kattan, L., Khan, A., 2011. A multinomial logit model
of pedestrian-vehicle crash severity. International Journal of
Sustainable Transportation 5, 233--249.
doi:\href{https://doi.org/10.1080/15568318.2010.497547}{10.1080/15568318.2010.497547}

\leavevmode\hypertarget{ref-Torrao2014modeling}{}%
Torrao, G.A., Coelho, M.C., Rouphail, N.M., 2014. Modeling the impact of
subject and opponent vehicles on crash severity in two-vehicle
collisions. Transportation Research Record 53--64.
doi:\href{https://doi.org/10.3141/2432-07}{10.3141/2432-07}

\leavevmode\hypertarget{ref-Train2009discrete}{}%
Train, K., 2009. Discrete choice methods with simulation, 2nd Edition.
ed. Cambridge University Press, Cambridge.

\leavevmode\hypertarget{ref-Wang2015copula}{}%
Wang, K., Yasmin, S., Konduri, K.C., Eluru, N., Ivan, J.N., 2015.
Copula-based joint model of injury severity and vehicle damage in
two-vehicle crashes. Transportation Research Record 158--166.
doi:\href{https://doi.org/10.3141/2514-17}{10.3141/2514-17}

\leavevmode\hypertarget{ref-Wang2005use}{}%
Wang, X.K., Kockelman, K.M., 2005. Use of heteroscedastic ordered logit
model to study severity of occupant injury - distinguishing effects of
vehicle weight and type, in: Statistical Methods; Highway Safety Data,
Analysis, and Evaluation; Occupant Protection; Systematic Reviews and
Meta-Analysis, Transportation Research Record. pp. 195--204.

\leavevmode\hypertarget{ref-White1972effects}{}%
White, S., Clayton, S., 1972. Some effects of alcohol, age of driver,
and estimated speed on the likelihood of driver injury. Accident
Analysis \& Prevention 4.

\leavevmode\hypertarget{ref-Wu2014mixed}{}%
Wu, Q., Chen, F., Zhang, G.H., Liu, X.Y.C., Wang, H., Bogus, S.M., 2014.
Mixed logit model-based driver injury severity investigations in single-
and multi-vehicle crashes on rural two-lane highways. Accident Analysis
and Prevention 72, 105--115.
doi:\href{https://doi.org/10.1016/j.aap.2014.06.014}{10.1016/j.aap.2014.06.014}

\leavevmode\hypertarget{ref-Yasmin2013evaluating}{}%
Yasmin, S., Eluru, N., 2013. Evaluating alternate discrete outcome
frameworks for modeling crash injury severity. Accident Analysis \&
Prevention 59, 506--521.
doi:\href{https://doi.org/https://doi.org/10.1016/j.aap.2013.06.040}{https://doi.org/10.1016/j.aap.2013.06.040}

\end{document}